### 2025-05-23 draftFiguresPaper.Rmd
adding script for the first figures of the paper

---
title: "draftFiguresPaper"
author: "Camille"
date: "2025-05-23"
output: html_document
---

```{r}
library(tidyverse)
"%ni%" = Negate("%in%")

datasetClassification <- read.csv("output/dataClassifToPlot.csv")

datasetCountries <- read.csv("dataNew/allgpt-lit_w_countries.csv")

datasetClassificationCountries <- merge(datasetClassification, datasetCountries %>% 
                                          select(X, Year, country) %>% 
                                          rename(ID = X),
                                        by = "ID") %>% 
  select(-reader)

datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)

table(datasetClassificationCountries$country)

# 1) split into one row per country
datasetClassificationCountries_long <- datasetClassificationCountries %>%
  separate_rows(country, sep = ";\\s*") %>%
  mutate(countries = str_trim(country))

# fix country names 
library(countrycode)

# 1. Direct mapping
datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
  mutate(
    iso3 = countrycode(countries,
                       origin      = "country.name",
                       destination = "iso3c")
  ) %>% 
  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))

### add income groups to the dataset

incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)

datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)

# 2) get total papers by country
papers_by_country <- datasetClassificationCountries_long_clean %>%
  group_by(country, iso3, .drop = F) %>% 
  count(iso3, name = "n_papers") %>%
  arrange(desc(n_papers)) %>% 
  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)

papers_by_country %>%
  slice_max(n_papers, n = 20) %>%                     # top 30
  mutate(
    # 1) reorder by paper count (so bars go largest→smallest),
    country = fct_reorder(country, n_papers),
    # 2) then shove "Unknown" to the very end
    country = fct_relevel(country, "Unknown", after = 0)
  ) %>%
  ggplot(aes(x = country, y = n_papers, fill = Income.group)) +
    geom_col() +
    coord_flip() +
    scale_fill_discrete(na.value = 'grey')+
    labs(
      x = "Country",
      y = "Number of papers",
      title = "Top 20 Countries by Paper Count"
    ) +
    theme_minimal()

#### make a map ####

### getting the data together
library(rnaturalearth)
library(terra)
library(tidyterra)

world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))

world_countries$iso_a3 <- ifelse(world_countries$sovereignt == "Somaliland", "SOM", world_countries$iso_a3)

world_countries_merged <- merge(world_countries, papers_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)

world_countries_merged$n_papers <- ifelse(is.na(world_countries_merged$n_papers), 0, world_countries_merged$n_papers)

my_breaks = c(1, 5, 20, 50, 150)

map_plot_reference <- ggplot() +
  geom_spatvector(data = world_countries_merged, aes(fill = n_papers)) +
  scale_fill_gradient(name = "nb of papers", trans = "log",
                        breaks = my_breaks, labels = my_breaks)+
  theme_void()

plot(map_plot_reference)

library(scales)          # for label formatting

# 1) Filter out Antarctica from your data:
world_no_antarctica <- 
  world_countries_merged %>%
  filter(name != "Antarctica")

# 2) Build the map with integer labels and no decimals:
map_plot_viridis_clean <- ggplot() +
  geom_spatvector(
    data = world_no_antarctica,
    aes(fill = n_papers),
    color = "grey70",
    size  = 0.1
  ) +
  scale_fill_viridis_c(
    name    = "Papers",
    # option  = "plasma",
    trans   = "log10",
    breaks  = my_breaks,
    labels  = label_number(accuracy = 1,    # round to whole numbers
                           big.mark = ",",
                           trim     = TRUE),
    na.value = "grey80"
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title    = "Global Publication Counts by Country",
    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5)
  )

# Print it
map_plot_viridis_clean

### add  monitor data to compare
# load the monitor dataset and subset it
monitors <- read.csv("dataNew/mergedMonitors_v7.csv")

# keep only reference grade monitors
monitorsRef <- subset(monitors, isMonitor)

# keep only active (different datasets have different ways of dealing with that)
# airnow
# no subset because no precise info on when active

# openaq
monitorsRef$diffDays <- as.numeric(difftime(monitorsRef$lastUpdated, monitorsRef$firstUpdated, units = "days"))

monitorsRef <- subset(monitorsRef, !(diffDays < 30 & firstUpdated < ymd_hms("2025-02-01T00:00:00+00:00") & database == "OpenAQ"))

# waqi
monitorsRef <- subset(monitorsRef, !(occurenceWAQI < 2 & database == "WAQI"))

# WHO I am not sure if we should filter points that are only active for a year
table(monitorsRef$database)

# we probably want to check if the results are consistent if removing embassy monitors
table(monitorsRef$isEmbassy)

monitorsRef_simple <- monitorsRef %>%
  # filter(!isEmbassy) %>% # here to filter embassies
  select(newID, longitude, latitude)

monitors_vect <- vect(monitorsRef_simple, geom=c("longitude", "latitude"), crs=crs(world_no_antarctica))

# 1. intersect: this returns only those points lying in (or on) each polygon,
#    with the polygon’s attributes copied over
pts_in_c <- intersect(monitors_vect, world_no_antarctica)

# 2. now simply tally by the country name
counts2 <- table(pts_in_c$iso_a3)

print(counts2)

# aggregate the intersected points by country NAME
agg <- aggregate(pts_in_c, by = "iso_a3", fun = length) %>% as.data.frame() %>% 
  rename(count = agg_newID) %>% 
  select(iso_a3, count)

# join back to the country layer
world_no_antarctica <- merge(world_no_antarctica, agg, by = "iso_a3", all.x = TRUE)

world_no_antarctica_df <- world_no_antarctica %>% as.data.frame()

world_no_antarctica_df$count <- ifelse(is.na(world_no_antarctica_df$count), 0, world_no_antarctica_df$count)

world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
  geom_point()+
  geom_smooth(method = "lm")+
  facet_wrap(~income_grp)

# world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
#   geom_point()+
#   geom_smooth(method = "lm")+
#   facet_wrap(~Income.group)

world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
  geom_point()+
  geom_smooth(method = "lm")

subset(world_no_antarctica_df, count < 10) %>% ggplot(aes(x = n_papers, y = count))+
  geom_point()+
  geom_smooth(method = "lm")

#### vs mean levels of pollution !!! to be done ####

##### categories over time #####
#

# 2) count papers per country-year
yearly <- datasetClassificationCountries_long_clean %>%
  count(Income.group, Year, name = "n_year")

# 3) ensure every country has an entry for every year in your span
years <- seq(min(datasetClassificationCountries_long_clean$Year),
             max(datasetClassificationCountries_long_clean$Year))

yearly_complete <- yearly %>%
  complete(Income.group, Year = years, fill = list(n_year = 0))

# 4) compute cumulative sum within each country
cum <- yearly_complete %>%
  group_by(Income.group) %>%
  arrange(Year, .by_group = TRUE) %>%
  mutate(n_cumulative = cumsum(n_year)) %>%
  ungroup()

cum$Income.group <- ifelse(is.na(cum$Income.group), "Unknown", cum$Income.group)

cum %>%
  ggplot(aes(x = Year, y = n_cumulative, col = Income.group)) +
    geom_line(linewidth = 2) +
    labs(
      x = "Year",
      y = "Cumulative # of Papers",
      title = "Growth of Papers by income group"
    ) +
    theme_minimal()

# then cumulative for different categories of papers

# 1) pivot your dummy columns into long form
df_cat_long <- datasetClassificationCountries_long_clean %>%
  pivot_longer(
    cols = c(Perceptions, Priority, Policy, Health, Behavior),
    names_to  = "category",
    values_to = "flag"
  ) %>%
  # keep only “paper belongs to this category”
  filter(flag == 1) %>%
  select(-flag)

# 2) count per year × income_group × category
yearly_cat <- df_cat_long %>%
  count(Year, Income.group, category, name = "n_papers")

# 3) fill zeros so lines don’t break
all_years <- seq(min(datasetClassificationCountries_long_clean$Year), max(datasetClassificationCountries_long_clean$Year))

yearly_cat <- yearly_cat %>%
  complete(
    Year, Income.group, category,
    fill = list(n_papers = 0)
  )

yearly_cat <- yearly_cat %>%
  group_by(Income.group, category) %>%
  arrange(Year, .by_group = TRUE) %>%
  mutate(n_cum = cumsum(n_papers)) %>%
  ungroup() %>% 
  # reorder so levels go from lowest to highest n_cum at the final Year
  mutate(category = fct_reorder2(category, Year, n_cum))

# 4) plot: one facet per category, colored by income group
ggplot(yearly_cat,
       aes(x = Year, y = n_cum, color = category)) +
  geom_line(linewidth = 2) +
  facet_wrap(~ Income.group, scales = "free_y") +
  theme_minimal()

# a) sum across incomes, then compute cumulative
yearly_cat_agg <- yearly_cat %>%
  group_by(Year, category) %>%
  summarise(n_year = sum(n_papers), .groups="drop") %>%
  arrange(category, Year) %>%
  group_by(category) %>%
  mutate(n_cum = cumsum(n_year)) %>%
  ungroup()

# b) plot one line per category
ggplot(yearly_cat_agg,
       aes(x = Year, y = n_cum, color = category)) +
  geom_line(linewidth = 2) +
  theme_minimal() +
  labs(
    x     = "Year",
    y     = "Cumulative # of Papers",
    color = "Category",
    title = "Cumulative Papers by Category (all incomes combined)"
  )

```

```{r linking nb of papers with pollution levels over space and time?}

```

### 2025-07-03 draftFiguresPaper.Rmd

@@
 
-papers_by_country %>%
+papers_by_country$Income.group <- factor(
+  papers_by_country$Income.group,
+  levels = c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+  ),
+  ordered = TRUE
+)
+
+pCountryPaperIncome <- papers_by_country %>%
   slice_max(n_papers, n = 20) %>%                     # top 30
@@
     theme_minimal()
+
+ggsave(plot = pCountryPaperIncome, filename = paste("plots/countryPapersIncome.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
 
@@
 map_plot_viridis_clean
+
+ggsave(plot = map_plot_viridis_clean, filename = paste("plots/mapPapersCount.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
 
@@
 
-cum %>%
+cum$Income.group <- factor(
+  cum$Income.group,
+  levels = c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+    # "Unknown"
+  ),
+  ordered = TRUE
+)
+
+ppapersIncome <- cum %>%
   ggplot(aes(x = Year, y = n_cumulative, col = Income.group)) +
-    geom_line(linewidth = 2) +
+    geom_line(linewidth = 1) +
     labs(
@@
     theme_minimal()
+
+ggsave(plot = ppapersIncome, filename = paste("plots/incomePaperYear.png", sep = ""),
+       dpi=600, width = 16, height = 8, units='cm')
 
@@
 
+yearly_cat$Income.group <- factor(
+  yearly_cat$Income.group,
+  levels = c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+    # "Unknown"
+  ),
+  ordered = TRUE
+)
+
 yearly_cat <- yearly_cat %>%
@@
 # 4) plot: one facet per category, colored by income group
-ggplot(yearly_cat,
+ppapersIncomeYear <- ggplot(yearly_cat,
        aes(x = Year, y = n_cum, color = category)) +
-  geom_line(linewidth = 2) +
+  geom_line(linewidth = 1) +
   facet_wrap(~ Income.group, scales = "free_y") +
   theme_minimal()
+
+ggsave(plot = ppapersIncomeYear, filename = paste("plots/incomeCategoryPapersYear.png", sep = ""),
+       dpi=600, width = 20, height = 12, units='cm')
 
@@
 # b) plot one line per category
-ggplot(yearly_cat_agg,
+ppapersYear <- ggplot(yearly_cat_agg,
        aes(x = Year, y = n_cum, color = category)) +
-  geom_line(linewidth = 2) +
+  geom_line(linewidth = 1) +
   theme_minimal() +
@@
 
+ggsave(plot = ppapersYear, filename = paste("plots/countryPapersYear.png", sep = ""),
+       dpi=600, width = 16, height = 8, units='cm')
+
 ```
@@
 ```
+
+```{r survey classification}
+
+library(tidyverse)
+"%ni%" = Negate("%in%")
+
+surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
+                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
+                                                 surveyClassification$surveyedCountries)
+
+# 1) split into one row per country
+surveyCountries_long <- surveyClassification %>%
+  separate_rows(surveyedCountries, sep = ";\\s*") %>%
+  mutate(countries = str_trim(surveyedCountries))
+
+# fix country names first manual then with iso3
+
+surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
+                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
+                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
+                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
+                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
+                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
+                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
+                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
+
+library(countrycode)
+
+# 1. Direct mapping
+surveyCountries_long_clean <- surveyCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
+incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
+
+surveyCountries_long_clean <- merge(surveyCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+# 2) get total surveys by country
+surveys_by_country <- surveyCountries_long_clean %>%
+  group_by(countries, iso3, .drop = F) %>% 
+  count(iso3, name = "n_surveys") %>%
+  arrange(desc(n_surveys)) %>% 
+  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+psurveyCountry <- surveys_by_country %>%
+  slice_max(n_surveys, n = 40) %>%                     # top 30
+  mutate(
+    # 1) reorder by paper count (so bars go largest→smallest),
+    country = fct_reorder(countries, n_surveys),
+    # 2) then shove "Unknown" to the very end
+    # country = fct_relevel(country, "Unknown", after = 0)
+  ) %>%
+  ggplot(aes(x = country, y = n_surveys, fill = Income.group)) +
+    geom_col() +
+    coord_flip() +
+    scale_fill_discrete(na.value = 'grey')+
+    labs(
+      x = "Country",
+      y = "Number of Surveys",
+      title = "Top 40 Countries by Survey Question Count"
+    ) +
+    theme_minimal()
+
+ggsave(plot = psurveyCountry, filename = paste("plots/countrySurveyIncome.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
+
+df_cat_long <- surveyCountries_long_clean %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  select(-flag)
+
+# 2) count per year × income_group × category
+overall_cat <- df_cat_long %>%
+  count(Income.group, category, name = "n_surveys")
+
+# 4) plot: one facet per category, colored by income group
+ggplot(overall_cat,
+       aes(y = n_surveys, x = category, fill = category)) +
+  geom_col() +
+  facet_wrap(~ Income.group, scales = "free_y") +
+  theme_minimal()
+
+## same but over time
+
+surveyCountries_long_clean_yearly <- surveyCountries_long_clean
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2018, surveyCountries_long_clean_yearly$year)
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)
+
+surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
+  # 1. build a single “years” string
+  mutate(
+    all_years = if_else(
+      is.na(repeatedMeasure) | repeatedMeasure == "",
+      as.character(year),
+      paste(year, repeatedMeasure, sep = ";")
+    ),
+    # drop every space character
+    all_years = str_remove_all(all_years, "\\s+")
+  ) %>%
+  separate_rows(all_years, sep = ";") %>%
+  mutate(year_num = as.integer(all_years)) %>%
+  select(-all_years, -repeatedMeasure)
+
+df_cat_long <- surveyCountries_long_clean_yearly_t %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  select(-flag)
+
+# 2) count per year × income_group × category
+yearly_cat <- df_cat_long %>%
+  count(year_num, Income.group, category, name = "n_surveys")
+
+# 3) fill zeros so lines don’t break
+all_years <- seq(min(surveyCountries_long_clean_yearly_t$year_num), max(surveyCountries_long_clean_yearly_t$year_num))
+
+yearly_cat_filled <- yearly_cat %>%
+  complete(
+    year_num = all_years,
+    Income.group, category,
+    fill = list(n_surveys = 0)
+  )
+
+yearly_cat_filled <- yearly_cat_filled %>%
+  group_by(Income.group, category) %>%
+  arrange(year_num, .by_group = TRUE) %>%
+  mutate(n_cum = cumsum(n_surveys)) %>%
+  ungroup() %>% 
+  # reorder so levels go from lowest to highest n_cum at the final Year
+  mutate(category = fct_reorder2(category, year_num, n_cum))
+
+# 4) plot: one facet per income group, colored by category
+ggplot(yearly_cat_filled,
+       aes(x = year_num, y = n_cum, color = category)) +
+  # geom_point()+
+  geom_line(linewidth = 2) +
+  facet_wrap(~ Income.group, scales = "free_y") +
+  theme_minimal()
+
+yearly_cat_filled$Income.group <- factor(
+  yearly_cat_filled$Income.group,
+  levels = c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+  ),
+  ordered = TRUE
+)
+
+pyear_survey <- ggplot(yearly_cat_filled,
+       aes(x = year_num, y = n_cum, color = category)) +
+  # geom_point()+
+  geom_line(linewidth = 1) +
+  facet_wrap(~ Income.group, scales = "free_y") +
+  theme_minimal()
+
+ggsave(plot = pyear_survey, filename = paste("plots/yearSurveyIncome.png", sep = ""),
+       dpi=600, width = 16, height = 8, units='cm')
+
+```
+
+```{r make a map for the survey data}
+
+#### make a map ####
+
+### getting the data together
+library(rnaturalearth)
+library(terra)
+library(tidyterra)
+library(tidyverse)
+
+"%ni%" = Negate("%in%")
+
+surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
+                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
+                                                 surveyClassification$surveyedCountries)
+
+# 1) split into one row per country
+surveyCountries_long <- surveyClassification %>%
+  separate_rows(surveyedCountries, sep = ";\\s*") %>%
+  mutate(countries = str_trim(surveyedCountries))
+
+# fix country names first manual then with iso3
+
+surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
+                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
+                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
+                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
+                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
+                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
+                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
+                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
+
+library(countrycode)
+
+# 1. Direct mapping
+surveyCountries_long_clean <- surveyCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
+incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
+
+surveyCountries_long_clean <- merge(surveyCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+# 2) get total surveys by country
+surveys_by_country <- surveyCountries_long_clean %>%
+  group_by(countries, iso3, .drop = F) %>% 
+  count(iso3, name = "n_surveys") %>%
+  arrange(desc(n_surveys)) %>% 
+  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))
+
+world_countries$iso_a3 <- ifelse(world_countries$sovereignt == "Somaliland", "SOM", world_countries$iso_a3)
+
+world_countries_merged <- merge(world_countries, surveys_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)
+
+world_countries_merged$n_surveys <- ifelse(is.na(world_countries_merged$n_surveys), 0, world_countries_merged$n_surveys)
+
+my_breaks = c(1, 5, 10, 20)
+
+map_plot_reference <- ggplot() +
+  geom_spatvector(data = world_countries_merged, aes(fill = n_surveys)) +
+  scale_fill_gradient(name = "nb of surveys", trans = "log",
+                        breaks = my_breaks, labels = my_breaks)+
+  theme_void()
+
+plot(map_plot_reference)
+
+library(scales)          # for label formatting
+
+# 1) Filter out Antarctica from your data:
+world_no_antarctica <- 
+  world_countries_merged %>%
+  filter(name != "Antarctica")
+
+# 2) Build the map with integer labels and no decimals:
+map_plot_viridis_clean_survey <- ggplot() +
+  geom_spatvector(
+    data = world_no_antarctica,
+    aes(fill = n_surveys),
+    color = "grey70",
+    size  = 0.1
+  ) +
+  scale_fill_viridis_c(
+    name    = "Papers",
+    # option  = "plasma",
+    trans   = "log10",
+    breaks  = my_breaks,
+    labels  = label_number(accuracy = 1,    # round to whole numbers
+                           big.mark = ",",
+                           trim     = TRUE),
+    na.value = "grey80"
+  ) +
+  coord_sf(expand = FALSE) +
+  labs(
+    title    = "Global Survey Question Counts by Country",
+    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
+  ) +
+  theme_void(base_size = 14) +
+  theme(
+    legend.position  = "bottom",
+    legend.key.width = unit(2, "cm"),
+    plot.title       = element_text(face = "bold", hjust = 0.5),
+    plot.subtitle    = element_text(hjust = 0.5)
+  )
+
+# Print it
+map_plot_viridis_clean_survey
+
+ggsave(plot = map_plot_viridis_clean_survey, filename = paste("plots/mapSurveyCount.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
+
+```

### 2025-07-03 draftFiguresPaper.Rmd
fixing plots and adding map

@@
 
+surveys_by_country$Income.group <- factor(
+  surveys_by_country$Income.group,
+  levels = c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+  ),
+  ordered = TRUE
+)
+
 psurveyCountry <- surveys_by_country %>%
@@
 
+yearly_cat_filled$category <- factor(
+  yearly_cat_filled$category,
+  levels = c(
+    "Perceptions",
+    "Health",
+    "Behavior",
+    "Policy",
+    "Priority"
+  ),
+  ordered = TRUE
+)
+
 pyear_survey <- ggplot(yearly_cat_filled,
@@
   scale_fill_viridis_c(
-    name    = "Papers",
+    name    = "Survey Questions",
     # option  = "plasma",

### 2025-08-05 comparisonClassificationThemesPapersv2.R
housekeeping

@@ -5,7 +5,7 @@ datasetHand <- read.csv("data/classificationByHand3.csv", sep = ";")
 datasetGPT <- read.csv("dataNew/classificationChatGPT_o4mini_final.csv")
 
 #### exclude papers that are excluded from the first script ####
-datasetGPTinclusion <- read.csv("dataNew/inclusionChatGPT_o4mini_fi nal.csv")
+datasetGPTinclusion <- read.csv("dataNew/inclusionChatGPT_o4mini_final.csv")
 
 getmode <- function(v) {
   v <- na.omit(v)

### 2025-08-05 comparisonInclusionGPTs.R

@@ -2,6 +2,24 @@ library(tidyverse)
 
 datasetGPTclassification <- read.csv("dataNew/classificationChatGPT_O4mini_final.csv")
 
+
+## identify papers that have ventilation focus
+
+# pattern: whole-word "ventilation", case-insensitive
+pattern <- "\\bventilation\\b"
+
+# logical vector of rows that match
+has_ventilation <- grepl(pattern, datasetGPTclassification$message_content, ignore.case = TRUE)
+
+# 1a) get row numbers
+row_nums <- which(has_ventilation)
+
+# 1b) subset the data.frame
+df_ventilation <- datasetGPTclassification[has_ventilation, ]
+
+#### end of identification
+
+
 cols <- c("Perceptions", "Behavior", "Policy", "Health", "Priority")
 
 # datasetGPTclassification <- subset(datasetGPTclassification, message_content != "")

### 2025-08-05 qualityAnalysis/countryQA.R

@@ -1,0 +1,41 @@
+library(tidyverse)
+
+classificationGPT <- read.csv("dataNew/allgpt-lit_w_countries.csv")
+
+classificationResearcher <- read.csv("dataNew/manualCodingCountry.csv")
+
+classificationResearcher$countries <- ifelse(classificationResearcher$countries %in% "Shouldn't be included?" | is.na(classificationResearcher$countries), "None", classificationResearcher$countries)
+
+classificationResearcher$countries <- ifelse(classificationResearcher$countries %in% "South Korea", "Korea, Republic of", 
+                                             ifelse(classificationResearcher$countries %in% "Taiwan", "Taiwan, Province of China", classificationResearcher$countries))
+
+idMatching <-  read.csv("output/tableIDmatching.csv")
+
+classificationGPT_toMatch <- merge(classificationGPT, idMatching %>% 
+                                     rename(TitleMatch = Title), by.x = "X", by.y = "IDgpt")
+
+mergedClassifs <- merge(classificationGPT_toMatch, classificationResearcher %>% 
+                          select(ID, countries) %>% 
+                          rename(countryHand = countries)
+                        , by.x = "IDhand", by.y = "ID")
+
+split_norm <- function(x) {
+  x %>% 
+    str_split("\\s*[;,]\\s*") %>% 
+    unlist() %>% 
+    str_trim() %>% 
+    discard(~ .x == "")
+}
+
+QAcountry <- mergedClassifs %>% 
+  mutate(
+    set1 = map(country, split_norm),
+    set2 = map(countryHand, split_norm),
+    inter = map2_int(set1, set2, ~ length(intersect(.x, .y))),
+    uni   = map2_int(set1, set2, ~ length(union(.x, .y))),
+    similarity = if_else(uni > 0, inter / uni, NA_real_)
+    )
+
+table(QAcountry$similarity)
+
+test1 <- read.csv("dataNew/datasetToClassifyGPT.csv")

### 2025-08-05 robustnessDraft.Rmd
robustness but very very messy next step is to rerun everything with a cleaner structure

--- 2025-05-23 draftFiguresPaper.Rmd
@@
 ---
-title: "draftFiguresPaper"
-author: "Camille"
-date: "2025-05-23"
+title: "robustnessDraft"
 output: html_document
 ---
+
+```{r identfying papers that are not in social science journals}
+
+litReviewSub <- read.csv("dataNew/combined_lit-socialscience.csv")
+litReview <- read.csv("dataNew/combined_lit.csv")
+
+clean_title <- function(x) {
+  x %>%
+    stri_trans_general("Latin-ASCII") %>%  # turn “é” → “e”
+    tolower() %>%                          # lowercase
+    gsub("[^[:alnum:] ]+", "", .) %>%      # drop punctuation
+    trimws()                               # trim whitespace
+}
+
+library(stringi)
+library(fuzzyjoin)
+
+litReviewSub <- litReviewSub %>%
+  select(Title) %>% 
+  mutate(title_clean = clean_title(Title))
+
+litReviewSub <- litReviewSub %>%
+  group_by(title_clean) %>% 
+  summarize(title_clean = first(title_clean))
+
+datasetID <- read.csv("dataNew/datasetToClassifyGPT.csv")
+
+datasetID <- datasetID %>%
+  mutate(title_clean = clean_title(Title))
+
+litReviewMerged2 <- stringdist_left_join(
+  litReviewSub, datasetID,
+  by = "title_clean",
+  method = "lv",        # Levenshtein
+  max_dist = 15,         # tweak this
+  distance_col = "dist" # keep the distance
+) %>%
+  arrange(-dist) %>%          # largest distances first
+  group_by(ID) %>% 
+  summarize(ID = first(ID),
+            Title = first(Title))
+
+write.csv(litReviewMerged2, "output/robustnessSocialScience.csv", row.names = F)
+
+```
+
+```{r wasting time matching something that should not be compliated}
+
+datasetClassification <- read.csv("output/dataClassifToPlot.csv")
+
+test <- read.csv("dataNew/allgpt-lit_w_countries.csv") %>% 
+  select(X , Title) %>% 
+  rename(ID = X)
+
+datasetClassificationMerged <- merge(datasetClassification, test, by = "ID")
+
+write.csv(datasetClassificationMerged, "output/dataClassifToPlotTemporaryRobust.csv", row.names = F)
+
+```
 
@@
 
+########### ALL BELOW TO DELETE AND ADAPT WHEN CLEAN CODE..... #########
+datasetClassification <- read.csv("output/dataClassifToPlotTemporaryRobust.csv")
+
+clean_title <- function(x) {
+  x %>%
+    stri_trans_general("Latin-ASCII") %>%  # turn “é” → “e”
+    tolower() %>%                          # lowercase
+    gsub("[^[:alnum:] ]+", "", .) %>%      # drop punctuation
+    trimws()                               # trim whitespace
+}
+
+library(stringi)
+library(fuzzyjoin)
+
+datasetClassificationR <- datasetClassification %>%
+  mutate(title_clean = clean_title(Title))
+
+robustnessSubset <- read.csv("output/robustnessSocialScience.csv")
+
+robustnessSubset <- robustnessSubset %>%
+  filter(!is.na(Title)) %>% 
+  mutate(title_clean = clean_title(Title)) %>% 
+  select(title_clean)
+
+datasetClassificationR2 <- stringdist_left_join(
+  datasetClassificationR, robustnessSubset,
+  by = "title_clean",
+  method = "lv",        # Levenshtein
+  max_dist = 15,         # tweak this
+  distance_col = "dist" # keep the distance
+) %>%
+  filter(!is.na(title_clean.y)) %>% 
+  arrange(-dist) %>%          # largest distances first
+  group_by(ID) %>% 
+  summarize(ID = first(ID),
+            Title = first(Title))
+
+datasetClassificationRobustIDs <- unique(datasetClassificationR2$ID)
+
+###### ALL ABOVE TO DELETE WHEN CLEAN CODE.... ######
 datasetClassification <- read.csv("output/dataClassifToPlot.csv")
-
 datasetCountries <- read.csv("dataNew/allgpt-lit_w_countries.csv")
 
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries %>% 
+datasetClassificationR <- subset(datasetClassification, ID %in% datasetClassificationRobustIDs)
+
+datasetClassificationCountries <- merge(datasetClassificationR, datasetCountries %>% 
                                           select(X, Year, country) %>% 
@@
     theme_minimal()
+
+ggsave(plot = pCountryPaperIncome, filename = paste("robustness/plots/countryPapersIncome.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
 
@@
 map_plot_viridis_clean
+
+ggsave(plot = map_plot_viridis_clean, filename = paste("robustness/plots/mapPapersCount.png", sep = ""),
+       dpi=600, width = 16, height = 12, units='cm')
 
@@
     theme_minimal()
+
+ggsave(plot = ppapersIncome, filename = paste("robustness/plots/incomePaperYear.png", sep = ""),
+       dpi=600, width = 16, height = 8, units='cm')
 
@@
 # 4) plot: one facet per category, colored by income group
-ggplot(yearly_cat,
+ppapersIncomeYear <- ggplot(yearly_cat,
        aes(x = Year, y = n_cum, color = category)) +
-  geom_line(linewidth = 2) +
+  geom_line(linewidth = 1) +
   facet_wrap(~ Income.group, scales = "free_y") +
   theme_minimal()
+
+ggsave(plot = ppapersIncomeYear, filename = paste("robustness/plots/incomeCategoryPapersYear.png", sep = ""),
+       dpi=600, width = 20, height = 12, units='cm')
 
@@
 
+ggsave(plot = ppapersYear, filename = paste("robustness/plots/countryPapersYear.png", sep = ""),
+       dpi=600, width = 16, height = 8, units='cm')
+
 ```

### 2025-08-05 qualityAnalysis/countryQA.R

@@ -37,5 +37,3 @@ QAcountry <- mergedClassifs %>%
     )
 
 table(QAcountry$similarity)
-
-test1 <- read.csv("dataNew/datasetToClassifyGPT.csv")

### 2025-08-06 qualityAnalysis/countryQA.R
adding unknownCountries

@@ -36,4 +36,4 @@ QAcountry <- mergedClassifs %>%
     similarity = if_else(uni > 0, inter / uni, NA_real_)
     )
 
-table(QAcountry$similarity)
+table(QAcountry$similarity) # almost perfect match

### 2025-09-10 qualityAnalysis/unknownCountries.R
manually classified missing countries next step will be to run the analysis again with the full dataset.

@@ -3,7 +3,11 @@ library(tidyverse)
 
 datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv") # to know which countries are excluded
 
-IDsubset <- unique(datasetClassification$ID)
+# for now remove studies that were classified as none of the categories during classification
+datasetClassificationSub <- subset(datasetClassification, None != 1)
+
+datasetClassificationSub 
+IDsubset <- unique(datasetClassificationSub$ID)
 
 # datasetCountries <- read.csv("LLMclassif/countriesChatGPT_5mini_final.csv")
 datasetCountries <- read.csv("LLMclassif/countriesChatGPT_o4mini_final.csv")

### 2025-09-11 draftFiguresPaper.Rmd
finalDatasetFigures lets discuss what gets included in the paper, the color scale, and whether we compare nb of papers and pollution levels

@@
 
-datasetClassification <- read.csv("output/dataClassifToPlot.csv")
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
 
-datasetCountries <- read.csv("dataNew/allgpt-lit_w_countries.csv")
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
 
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries %>% 
-                                          select(X, Year, country) %>% 
-                                          rename(ID = X),
+datasetCountries <- datasetCountries %>% 
+  select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
+
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
                                         by = "ID") %>% 
@@
 datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ";\\s*") %>%
+  separate_rows(country, sep = ",\\s*") %>%
   mutate(countries = str_trim(country))
 
+datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
+                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
+                                                               datasetClassificationCountries_long$countries))
 # fix country names 
@@
     # 1) reorder by paper count (so bars go largest→smallest),
-    country = fct_reorder(country, n_papers),
-    # 2) then shove "Unknown" to the very end
-    country = fct_relevel(country, "Unknown", after = 0)
+    country = fct_reorder(country, n_papers)
   ) %>%
@@
 
-### add  monitor data to compare
-# load the monitor dataset and subset it
-monitors <- read.csv("dataNew/mergedMonitors_v7.csv")
-
-# keep only reference grade monitors
-monitorsRef <- subset(monitors, isMonitor)
-
-# keep only active (different datasets have different ways of dealing with that)
-# airnow
-# no subset because no precise info on when active
-
-# openaq
-monitorsRef$diffDays <- as.numeric(difftime(monitorsRef$lastUpdated, monitorsRef$firstUpdated, units = "days"))
-
-monitorsRef <- subset(monitorsRef, !(diffDays < 30 & firstUpdated < ymd_hms("2025-02-01T00:00:00+00:00") & database == "OpenAQ"))
-
-# waqi
-monitorsRef <- subset(monitorsRef, !(occurenceWAQI < 2 & database == "WAQI"))
-
-# WHO I am not sure if we should filter points that are only active for a year
-table(monitorsRef$database)
-
-# we probably want to check if the results are consistent if removing embassy monitors
-table(monitorsRef$isEmbassy)
-
-monitorsRef_simple <- monitorsRef %>%
-  # filter(!isEmbassy) %>% # here to filter embassies
-  select(newID, longitude, latitude)
-
-monitors_vect <- vect(monitorsRef_simple, geom=c("longitude", "latitude"), crs=crs(world_no_antarctica))
-
-# 1. intersect: this returns only those points lying in (or on) each polygon,
-#    with the polygon’s attributes copied over
-pts_in_c <- intersect(monitors_vect, world_no_antarctica)
-
-# 2. now simply tally by the country name
-counts2 <- table(pts_in_c$iso_a3)
-
-print(counts2)
-
-# aggregate the intersected points by country NAME
-agg <- aggregate(pts_in_c, by = "iso_a3", fun = length) %>% as.data.frame() %>% 
-  rename(count = agg_newID) %>% 
-  select(iso_a3, count)
-
-# join back to the country layer
-world_no_antarctica <- merge(world_no_antarctica, agg, by = "iso_a3", all.x = TRUE)
-
-world_no_antarctica_df <- world_no_antarctica %>% as.data.frame()
-
-world_no_antarctica_df$count <- ifelse(is.na(world_no_antarctica_df$count), 0, world_no_antarctica_df$count)
-
-world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
-  geom_point()+
-  geom_smooth(method = "lm")+
-  facet_wrap(~income_grp)
-
-# world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
-#   geom_point()+
-#   geom_smooth(method = "lm")+
-#   facet_wrap(~Income.group)
-
-world_no_antarctica_df %>% ggplot(aes(x = n_papers, y = count))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-subset(world_no_antarctica_df, count < 10) %>% ggplot(aes(x = n_papers, y = count))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-#### vs mean levels of pollution !!! to be done ####
-
 ##### categories over time #####
 #
+
+# data with ID year 
+datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
+  select(ID, Year)
+
+datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
+                                                   datasetYear,
+                                                   by = "ID", all.x = T, all.y = F)
 

### 2025-09-11 qualityAnalysis/unknownCountries.R

@@ -134,13 +134,19 @@ write.csv(unknownCountriesSave, "output/unknownCountryo4mini.csv", row.names = F
 ######### after manually coding the countries:
 
 # load manually coded remaining papers
-manualCoded <- read.csv("dataNew/unknownCountryo4mini_CF.csv")
+manualCoded <- read.csv("dataNew/unknownCountryo4mini_CF.csv", sep = ";")
+manualCoded <- subset(manualCoded, !is.na(countryCF))  
 
 # merge and save for final classif
 
 datasetCountriesPartial_f <- merge(datasetCountriesPartial, manualCoded %>%
-                                     select(ID, manualCountry),
+                                     select(ID, countryCF),
                                    all.x = T, all.y = F, by = "ID")
+
+datasetCountriesPartial_f$countryPartialF <- ifelse(datasetCountriesPartial_f$countryPartial == "Unknown", datasetCountriesPartial_f$countryCF,
+                                                    datasetCountriesPartial_f$countryPartial)
+
+datasetCountriesPartial_f <- subset(datasetCountriesPartial_f, !is.na(countryPartialF))
 
 write.csv(datasetCountriesPartial_f, "output/fullClassifCountryo4mini.csv", row.names = F) # this should be used for final plotting
 

### 2025-09-11 draftFiguresPaper.Rmd
cleanup & robustness only social science journals

@@
 
-ppapersIncome <- cum %>%
+ppapersIncome <- subset(cum, !is.na(Income.group)) %>% 
   ggplot(aes(x = Year, y = n_cumulative, col = Income.group)) +
@@
 # 4) plot: one facet per category, colored by income group
-ppapersIncomeYear <- ggplot(yearly_cat,
+ppapersIncomeYear <- ggplot(subset(yearly_cat, !is.na(Income.group)),
        aes(x = Year, y = n_cum, color = category)) +

### 2025-09-11 robustnessDraft.Rmd
adding unknownCountries
cleanup & robustness only social science journals

@@
 
+library(tidyverse)
+library(stringi)
+library(fuzzyjoin)
+
+### !!!!! ihi something weird about papers that have a translation
+
 litReviewSub <- read.csv("dataNew/combined_lit-socialscience.csv")
-litReview <- read.csv("dataNew/combined_lit.csv")
 
@@
 }
-
-library(stringi)
-library(fuzzyjoin)
 
@@
 
-```{r wasting time matching something that should not be compliated}
-
-datasetClassification <- read.csv("output/dataClassifToPlot.csv")
-
-test <- read.csv("dataNew/allgpt-lit_w_countries.csv") %>% 
-  select(X , Title) %>% 
-  rename(ID = X)
-
-datasetClassificationMerged <- merge(datasetClassification, test, by = "ID")
-
-write.csv(datasetClassificationMerged, "output/dataClassifToPlotTemporaryRobust.csv", row.names = F)
-
-```
+# ```{r wasting time matching something that should not be compliated}
+# 
+# datasetClassification <- read.csv("output/dataClassifToPlot.csv")
+# 
+# test <- read.csv("dataNew/allgpt-lit_w_countries.csv") %>% 
+#   select(X , Title) %>% 
+#   rename(ID = X)
+# 
+# datasetClassificationMerged <- merge(datasetClassification, test, by = "ID")
+# 
+# write.csv(datasetClassificationMerged, "output/dataClassifToPlotTemporaryRobust.csv", row.names = F)
+# 
+# ```
 
@@
 
-########### ALL BELOW TO DELETE AND ADAPT WHEN CLEAN CODE..... #########
-datasetClassification <- read.csv("output/dataClassifToPlotTemporaryRobust.csv")
+dataToSubset <- read.csv("output/robustnessSocialScience.csv")
 
-clean_title <- function(x) {
-  x %>%
-    stri_trans_general("Latin-ASCII") %>%  # turn “é” → “e”
-    tolower() %>%                          # lowercase
-    gsub("[^[:alnum:] ]+", "", .) %>%      # drop punctuation
-    trimws()                               # trim whitespace
-}
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
 
-library(stringi)
-library(fuzzyjoin)
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
 
-datasetClassificationR <- datasetClassification %>%
-  mutate(title_clean = clean_title(Title))
+datasetCountries <- datasetCountries %>% 
+  select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
 
-robustnessSubset <- read.csv("output/robustnessSocialScience.csv")
-
-robustnessSubset <- robustnessSubset %>%
-  filter(!is.na(Title)) %>% 
-  mutate(title_clean = clean_title(Title)) %>% 
-  select(title_clean)
-
-datasetClassificationR2 <- stringdist_left_join(
-  datasetClassificationR, robustnessSubset,
-  by = "title_clean",
-  method = "lv",        # Levenshtein
-  max_dist = 15,         # tweak this
-  distance_col = "dist" # keep the distance
-) %>%
-  filter(!is.na(title_clean.y)) %>% 
-  arrange(-dist) %>%          # largest distances first
-  group_by(ID) %>% 
-  summarize(ID = first(ID),
-            Title = first(Title))
-
-datasetClassificationRobustIDs <- unique(datasetClassificationR2$ID)
-
-###### ALL ABOVE TO DELETE WHEN CLEAN CODE.... ######
-datasetClassification <- read.csv("output/dataClassifToPlot.csv")
-datasetCountries <- read.csv("dataNew/allgpt-lit_w_countries.csv")
-
-datasetClassificationR <- subset(datasetClassification, ID %in% datasetClassificationRobustIDs)
-
-datasetClassificationCountries <- merge(datasetClassificationR, datasetCountries %>% 
-                                          select(X, Year, country) %>% 
-                                          rename(ID = X),
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
                                         by = "ID") %>% 
@@
 
+# here is the step where we only keep papers in social science journals
+datasetClassificationCountries <- subset(datasetClassificationCountries, ID %in% unique(dataToSubset$ID))
+
 table(datasetClassificationCountries$country)
@@
 
-ppapersIncome <- cum %>%
+ppapersIncome <- subset(cum, !is.na(Income.group)) %>%
   ggplot(aes(x = Year, y = n_cumulative, col = Income.group)) +
@@
 ```
-
-```{r linking nb of papers with pollution levels over space and time?}
-
-```

### 2025-09-11 comparingInclusionBatch.R

@@ -1,93 +1,0 @@
-library(tidyverse)
-
-datasetGPTBatch1 <- read.csv("data/inclusionChatGPT_test2.csv", sep = ",") %>% 
-  rename(ID = custom_id,
-         messageBatch = message_content,
-         inclusionBatch = inclusionChatGPTClean)
-
-datasetGPT <- read.csv("data/inclusionChatGPT1704.csv")
-
-datasetGPTmerge <- datasetGPT %>%
-  rename(messageOld = response,
-         inclusionOld = inclusionChatGPT) %>% 
-  select(ID, messageOld, inclusionOld)
-
-datasetGPTmerge$inclusionOld <- gsub(
-  "\\[|\\]|'",         # pattern: any literal "[", "]" or "'"
-  "",                  # replacement: nothing
-  datasetGPTmerge$inclusionOld
-)
-
-test <- merge(datasetGPTBatch1, datasetGPTmerge)
-
-#### Merging classifications to make comparisons
-
-datasetHand <- read.csv("data/classificationByHand3.csv", sep = ";")
-
-datasetGPTBatch1 <- read.csv("data/inclusionChatGPT_test2.csv", sep = ",") %>% 
-  rename(ID = custom_id,
-         messageBatch = message_content,
-         Inclusion = inclusionChatGPTClean) %>% 
-  select(ID, Inclusion) %>% 
-  mutate(reader = "ChatBatch")
-
-datasetGPT <- read.csv("data/inclusionChatGPT1704.csv")
-
-datasetGPTmerge <- datasetGPT %>%
-  rename(Inclusion = inclusionChatGPT) %>% 
-  select(ID, Inclusion) %>% 
-  mutate(reader = "ChatOld")
-
-datasetGPTmerge$Inclusion <- gsub(
-  "\\[|\\]|'",         # pattern: any literal "[", "]" or "'"
-  "",                  # replacement: nothing
-  datasetGPTmerge$Inclusion
-)
-
-GPTtogether <- rbind(datasetGPTBatch1, datasetGPTmerge)
-
-GPTtogether$Inclusion <- ifelse(GPTtogether$Inclusion == "included", 1, 0)
-
-datasetHand_fit <- datasetHand %>% 
-  select(ID, Inclusion, reader)
-
-dataCompare <- rbind(datasetHand_fit, GPTtogether)
-
-cols <- c("Inclusion")
-
-class_cols <- cols
-
[…]

### 2025-09-11 comparisonClassificationThemesPapers.R

@@ -1,279 +1,0 @@
-########## comparing classif
-library(tidyverse)
-
-datasetHand <- read.csv("data/classificationByHand3.csv", sep = ";")
-
-datasetGPT <- read.csv("data/classificationChatGPT_o4mini.csv")
-
-cols <- c("Perceptions", "Behavior", "Policy", "Health", "Priority")
-
-# cols <- c("inclusion", "Perceptions", "Behavior", "Policy", "Health", "Priority")
-
-datasetHand[cols] <- lapply(datasetHand[cols], function(x) ifelse(is.na(x) | x == "?" | x == " " | x == "", 0, x))
-
-datasetGPT <- subset(datasetGPT, message_content != "")
-
-datasetGPTsimple <- datasetGPT %>%
-  select(custom_id, choice_index, classificationChatGPTClean)
-
-datasetGPTsimple$classificationChatGPTClean <- ifelse(datasetGPTsimple$classificationChatGPTClean == "", "None", datasetGPTsimple$classificationChatGPTClean)
-
-getmode <- function(v) {
-  v <- na.omit(v)
-  uniqv <- unique(v)
-  uniqv[which.max(tabulate(match(v, uniqv)))]
-}
-
-colsGPT <- c("Perceptions", "Behaviors", "Policy", "Health", "Priority", "None")
-
-datasetGPTsimple_wide <- datasetGPTsimple %>%
-  group_by(custom_id, choice_index) %>% 
-  # Remove brackets and single quotes
-  mutate(categories = gsub("\\[|\\]|'", "", classificationChatGPTClean)) %>%
-  # Split the string into multiple rows
-  separate_rows(categories, sep = ",\\s*") %>%
-  mutate(categories = str_replace_all(categories, " ", "")) %>% 
-  # Create a dummy column
-  mutate(dummy = 1) %>%
-  # Pivot to a wide format, filling missing values with 0
-  tidyr::pivot_wider(names_from = categories, values_from = dummy, values_fill = 0) %>% 
-  group_by(custom_id) %>% 
-  summarise(
-    across(
-      all_of(colsGPT),   # or just `categories` if you’re unquoted
-      getmode,              # the function to apply
-      .names = "{.col}"     # leave the names unchanged; use "mode_{.col}" if you prefer
-    )
-  )
-
-datasetGPTsimple_wide_fit <- datasetGPTsimple_wide %>% 
-  rename(
[…]

### 2025-09-11 comparisonClassificationThemesPapersv2.R

@@ -1,363 +1,0 @@
-library(tidyverse)
-
-datasetHand <- read.csv("data/classificationByHand3.csv", sep = ";")
-
-datasetGPT <- read.csv("dataNew/classificationChatGPT_o4mini_final.csv")
-
-#### exclude papers that are excluded from the first script ####
-datasetGPTinclusion <- read.csv("dataNew/inclusionChatGPT_o4mini_final.csv")
-
-getmode <- function(v) {
-  v <- na.omit(v)
-  uniqv <- unique(v)
-  uniqv[which.max(tabulate(match(v, uniqv)))]
-}
-
-# similarly for Gini
-norm_gini <- function(n) {
-  p <- n / sum(n)
-  G <- 1 - sum(p^2)
-  G / (1 - 1/length(n))
-}
-
-datasetGPTsummary <- datasetGPTinclusion %>%
-  # filter(inclusionChatGPTClean %in% c("excluded", "included")) %>% 
-  group_by(custom_id) %>% 
-  summarize(inclusionChatGPTCleanMode = getmode(inclusionChatGPTClean),
-            countExclusion = sum(inclusionChatGPTClean == "excluded", na.rm = TRUE),
-            countInclusion = sum(inclusionChatGPTClean == "included", na.rm = TRUE),
-            countBlank = sum(inclusionChatGPTClean == "", na.rm=T))
-
-datasetGPTsummary$gini_norm <- mapply(
-  function(a, b, c) norm_gini(c(a, b, c)),
-  datasetGPTsummary$countExclusion,
-  datasetGPTsummary$countInclusion,
-  datasetGPTsummary$countBlank
-)
-
-#IDs to check manually for exclusion from inclusion data
-IDmanual2 <- unique(subset(datasetGPTsummary, gini_norm > 0.5)$custom_id)
-IDmanual3 <- subset(datasetGPTsummary, countBlank > 1)$custom_id
-
-save(IDmanual2, IDmanual3, file = "output/checkIDmanualInclusion.RData")
-
-# these are the uncertain ones, for now I remove here
-datasetGPTsummary <- subset(datasetGPTsummary, gini_norm < 0.5)
-datasetGPTsummary <- subset(datasetGPTsummary, countBlank < 2)
-
-datasetInclusion <- subset(datasetGPTsummary, countInclusion > 2)
-
-##### make sure the IDs match #####
-dataMatching <- read.csv("output/tableIDMatching.csv")
-
[…]

### 2025-09-11 comparisonInclusionGPTs.R

@@ -1,150 +1,0 @@
-library(tidyverse)
-
-datasetGPTclassification <- read.csv("dataNew/classificationChatGPT_O4mini_final.csv")
-
-
-## identify papers that have ventilation focus
-
-# pattern: whole-word "ventilation", case-insensitive
-pattern <- "\\bventilation\\b"
-
-# logical vector of rows that match
-has_ventilation <- grepl(pattern, datasetGPTclassification$message_content, ignore.case = TRUE)
-
-# 1a) get row numbers
-row_nums <- which(has_ventilation)
-
-# 1b) subset the data.frame
-df_ventilation <- datasetGPTclassification[has_ventilation, ]
-
-#### end of identification
-
-
-cols <- c("Perceptions", "Behavior", "Policy", "Health", "Priority")
-
-# datasetGPTclassification <- subset(datasetGPTclassification, message_content != "")
-
-datasetGPTsimple <- datasetGPTclassification %>%
-  select(custom_id, choice_index, classificationChatGPTClean)
-
-datasetGPTsimple$classificationChatGPTClean <- ifelse(datasetGPTsimple$classificationChatGPTClean == "", "Blank", datasetGPTsimple$classificationChatGPTClean)
-
-getmode <- function(v) {
-  v <- na.omit(v)
-  uniqv <- unique(v)
-  uniqv[which.max(tabulate(match(v, uniqv)))]
-}
-
-colsGPT <- c("Perceptions", "Behaviors", "Policy", "Health", "Priority", "None", "Blank")
-
-# find a way to measure uncertainty for each category
-
-# for sure, countBlanks, but then what, count the number for every category? Could work
-
-datasetGPTsimple_wide <- datasetGPTsimple %>%
-  group_by(custom_id, choice_index) %>% 
-  # Remove brackets and single quotes
-  mutate(categories = gsub("\\[|\\]|'", "", classificationChatGPTClean)) %>%
-  # Split the string into multiple rows
-  separate_rows(categories, sep = ",\\s*") %>%
-  mutate(categories = str_replace_all(categories, " ", "")) %>% 
-  # Create a dummy column
-  mutate(dummy = 1) %>%
-  # Pivot to a wide format, filling missing values with 0
-  tidyr::pivot_wider(names_from = categories, values_from = dummy, values_fill = 0) %>% 
[…]

### 2025-10-16 draftFiguresPaper.Rmd
summmary dataset to easily draft plots

@@
 
+```{r putting data together}
+
+library(tidyverse)
+"%ni%" = Negate("%in%")
+
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
+
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
+
+datasetCountries <- datasetCountries %>% 
+  select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
+
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
+                                        by = "ID") %>% 
+  select(-reader)
+
+datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
+
+table(datasetClassificationCountries$country)
+
+# 1) split into one row per country
+datasetClassificationCountries_long <- datasetClassificationCountries %>%
+  separate_rows(country, sep = ",\\s*") %>%
+  mutate(countries = str_trim(country))
+
+datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
+                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
+                                                               datasetClassificationCountries_long$countries))
+# fix country names 
+library(countrycode)
+
+# 1. Direct mapping
+datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+### add income groups to the dataset
+
+incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
+incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
+
+datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+# 2) get total papers by country
+papers_by_country <- datasetClassificationCountries_long_clean %>%
+  group_by(country, iso3, .drop = F) %>% 
+  count(iso3, name = "n_papers") %>%
+  arrange(desc(n_papers)) %>% 
+  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")
+
+datasetPollution_avg <- datasetPollution %>% 
+  filter(Year == 2019) %>% 
+  group_by(Region) %>% 
+  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
+
+datasetPollution_ctr <- datasetPollution_avg %>%
+  mutate(
+    iso3 = countrycode(Region,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(!is.na(iso3)) %>% 
+  rename(countryName = Region)
+
+papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)
+
+fullCountrylist <- unique(countrycode::codelist$iso3c)
+
+papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)
+
+iso3_un_members <- c(
+  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
+  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
+  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
+  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
+  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
+  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
+  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
+  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
+  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
+  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
+  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
+  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
+  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
+  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
+  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
+  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
+  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
+  "VUT","VEN","VNM","YEM","ZMB","ZWE"
+)
+
+papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
+  filter(!is.na(iso3))
+
+## adding survey nb
+
+surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
+                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
+                                                 surveyClassification$surveyedCountries)
+
+# 1) split into one row per country
+surveyCountries_long <- surveyClassification %>%
+  separate_rows(surveyedCountries, sep = ";\\s*") %>%
+  mutate(countries = str_trim(surveyedCountries))
+
+# fix country names first manual then with iso3
+
+surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
+                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
+                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
+                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
+                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
+                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
+                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
+                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
+
+# 1. Direct mapping
+surveyCountries_long_clean <- surveyCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+# 2) get total surveys by country
+surveys_by_country <- surveyCountries_long_clean %>%
+  group_by(iso3, .drop = F) %>% 
+  count(iso3, name = "n_surveys") %>%
+  arrange(desc(n_surveys))
+
+papers_by_country_pollution_filteredUN_survey <- merge(papers_by_country_pollution_filteredUN, surveys_by_country, by = "iso3", all.x = T, all.y = T)
+
+papers_by_country_pollution_filteredUN_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_filteredUN_survey$n_surveys), 0, papers_by_country_pollution_filteredUN_survey$n_surveys)
+
+population <- read.csv("dataNew/GHS_pop.csv", sep = ";")
+
+popCtr <- population %>%
+  group_by(GADM_ISO) %>% 
+  summarize(population2015 = sum(X2015))
+
+summaryDataset <- merge(papers_by_country_pollution_filteredUN_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)
+
+summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+                  labels = c("very low", "low", "medium", "high", "very high"), 
+                  right = FALSE)
+
+plot(summaryDataset %>% 
+       select(Income.group, pm25Avg, n_papers, n_surveys, population2015, categoryPollution))
+
+summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
+summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)
+
+write.csv(summaryDataset, "output/datasetPlots.csv", row.names = F)
+
+```
+
 ```{r linking nb of papers with pollution levels over space and time?}
 
+library(tidyverse)
+"%ni%" = Negate("%in%")
+
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
+
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
+
+datasetCountries <- datasetCountries %>% 
+  select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
+
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
+                                        by = "ID") %>% 
+  select(-reader)
+
+datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
+
+table(datasetClassificationCountries$country)
+
+# 1) split into one row per country
+datasetClassificationCountries_long <- datasetClassificationCountries %>%
+  separate_rows(country, sep = ",\\s*") %>%
+  mutate(countries = str_trim(country))
+
+datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
+                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
+                                                               datasetClassificationCountries_long$countries))
+# fix country names 
+library(countrycode)
+
+# 1. Direct mapping
+datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+### add income groups to the dataset
+
+incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
+incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
+
+datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+# 2) get total papers by country
+papers_by_country <- datasetClassificationCountries_long_clean %>%
+  group_by(country, iso3, .drop = F) %>% 
+  count(iso3, name = "n_papers") %>%
+  arrange(desc(n_papers)) %>% 
+  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")
+
+datasetPollution_avg <- datasetPollution %>% 
+  filter(Year == 2019) %>% 
+  group_by(Region) %>% 
+  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
+
+datasetPollution_ctr <- datasetPollution_avg %>%
+  mutate(
+    iso3 = countrycode(Region,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(!is.na(iso3))
+
+papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)
+
+fullCountrylist <- unique(countrycode::codelist$iso3c)
+
+papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)
+
+papers_by_country_pollution_filtered <- subset(papers_by_country_pollution, iso3 %in% fullCountrylist) %>% 
+  filter(!is.na(iso3))
+
+papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()
+
+papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth()
+
+papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+subset(papers_by_country_pollution_filtered, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filtered %>% 
+  filter(n_papers > 0) %>%
+  filter(iso3 != "CHN") %>% 
+  ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filtered %>% 
+  # filter(n_papers > 0) %>%
+  filter(iso3 != "CHN") %>% 
+  ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+iso3_un_members <- c(
+  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
+  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
+  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
+  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
+  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
+  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
+  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
+  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
+  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
+  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
+  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
+  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
+  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
+  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
+  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
+  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
+  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
+  "VUT","VEN","VNM","YEM","ZMB","ZWE"
+)
+
+papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
+  filter(!is.na(iso3))
+
+papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()
+
+papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth()
+
+papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
+  geom_point()+
+  geom_smooth(method = "lm")+
+  ylim(0,40)
+
+subset(papers_by_country_pollution_filteredUN, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filteredUN %>% 
+  filter(n_papers > 0) %>%
+  filter(iso3 != "CHN") %>% 
+  ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filteredUN %>% 
+  # filter(n_papers > 0) %>%
+  filter(iso3 != "CHN") %>% 
+  ggplot(aes(x = pm25Avg, y = n_papers))+
+  geom_point()+
+  geom_smooth(method = "lm")
+
+papers_by_country_pollution_filteredUN %>% ggplot(aes(x = n_papers > 0, y = pm25Avg))+
+  geom_boxplot()
+
+papers_by_country_pollution_filteredUN$categoryPollution <- cut(papers_by_country_pollution_filteredUN$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+                  labels = c("very low", "low", "medium", "high", "very high"), 
+                  right = FALSE)
+
+papers_by_country_pollution_filteredUN %>% 
+  # filter(iso3 != "CHN") %>%
+  ggplot(aes(x = categoryPollution, y = log(n_papers)))+
+  geom_boxplot()
+
+papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
+  mutate(paperBin = ifelse(n_papers > 0, 1, 0)) %>% 
+  group_by(categoryPollution) %>% 
+  summarize(proportionPaper = sum(paperBin, na.rm=T)/n(),
+            nbCountries = n())
+
+papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = proportionPaper))+
+  geom_boxplot()+
+  ylim(0,1)
+
+papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
+  group_by(categoryPollution) %>% 
+  summarize(papersPerCountry = sum(n_papers, na.rm=T)/n(),
+            nbCountries = n(),
+            medianPapersPerCountry = median(n_papers, na.rm=T))
+
+papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = medianPapersPerCountry))+
+  geom_boxplot()
+
+table(papers_by_country_pollution_filteredUN$categoryPollution)
+
+papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = papersPerCountry))+
+  geom_boxplot()
+
+TEST <- papers_by_country_pollution_filteredUN %>% 
+  filter(categoryPollution == "very high")
+
+# ylim(0,1)
 ```

### 2025-10-20 draftFiguresPaper.Rmd
add dataset
save datasets for Figure2

@@
                                                    by = "ID", all.x = T, all.y = F)
+
+## save this dataset for further analyses
+write.csv(datasetClassificationCountries_long_clean, "output/datasetClassificationCountries_long_clean.csv", row.names = F)
 
@@
 
+# add region from natural earth
+
+library(rnaturalearth)
+
+# .
+# .
+# .
+
 write.csv(summaryDataset, "output/datasetPlots.csv", row.names = F)
@@
 surveyCountries_long_clean <- merge(surveyCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+
+## save this dataset for further analyses
+write.csv(surveyCountries_long_clean, "output/datasetSurveyCountries_long_clean.csv", row.names = F)
 

### 2025-10-20 draftFiguretwo_Ella.Rmd
some options for Figure 2

---
title: "Figure two"
author: "Ella Henninger"
date: "2025-10-17"
output: html_document
---

```{r setup, include=FALSE}

## clean environment
rm(list = ls())

# Define which packages needed for analyses
p_needed <-
  c("knitr",
    "dplyr", 
    "stargazer",
    "ggplot2",
    "viridis",
    "haven",
    "MASS",
    "modelsummary",
    "tidyr",
    "data.table",
    "scales")

# Check which packages are already installed on your computer
packages <- rownames(installed.packages())

# Check which packages are not installed
p_to_install <- p_needed[!(p_needed %in% packages)]

if (length(p_to_install) > 0) {
  install.packages(p_to_install)
}
sapply(p_needed, require, character.only = TRUE)

# Set an option for the final document that can be produced from the .Rmd file.
knitr::opts_chunk$set(echo = TRUE)

## For replicability: session information 
session_info <- print(sessionInfo())
```

## Load the data

```{r data}

#-------------------------------------------------------------------------------
# Load the data
#-------------------------------------------------------------------------------
papers <- read.csv2("output/datasetClassificationCountries_long_clean.csv", 
                    header = T, sep = ",", dec = ".")

surveys <- read.csv2("output/datasetSurveyCountries_long_clean.csv", 
                     header = T, sep = ",", dec = ".")

```

## Clean the data

```{r clean}

#-------------------------------------------------------------------------------
# consistent country names 
#-------------------------------------------------------------------------------

## paper
papers <- papers %>%
  mutate(country_clean = recode(country,
    # Turkey variants
    "Turkey" = "Turkyie",
    "Türkiye" = "Turkyie",

    # UK variants
    "UK" = "United Kingdom",
    "England" = "United Kingdom",
    "Wales" = "United Kingdom",

    # Korea variants
    "Korea" = "South Korea",

    # Netherlands variants
    "The Netherlands" = "Netherlands",

    # Czechia variants
    "Czech Republic" = "Czechia",

    # US variants
    "USA" = "United States",

    .default = country  # keep as is otherwise
  ))

papers <- papers[which(papers$country_clean != "exclude"),]
papers <- papers[which(papers$country_clean != "no info"),]

## survey
surveys <- surveys %>%
  mutate(country_clean = recode(countries,
    # Turkey variants
    "Turkey" = "Turkyie",
    "Türkiye" = "Turkyie",

    # UK variants
    "UK" = "United Kingdom",
    "England" = "United Kingdom",
    "Wales" = "United Kingdom",

    # Korea variants
    "Korea" = "South Korea",

    # Netherlands variants
    "The Netherlands" = "Netherlands",

    # Czechia variants
    "Czech Republic" = "Czechia",

    # US variants
    "USA" = "United States",

    .default = countries  # keep as is otherwise
  ))

#-------------------------------------------------------------------------------
# relevant columns
#-------------------------------------------------------------------------------
papers <- papers[,c("ID", "country_clean",
                    "Income.group", "Region", 
                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "None")]

surveys <- surveys[,c("id", "country_clean", "year", "Income.group", "Region", 
                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "Inclusion")]

```

## Visualizations full data

```{r visualise}

#-------------------------------------------------------------------------------
# papers
#-------------------------------------------------------------------------------

# reshape to long format and count topics
papers_long <- papers %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
    names_to = "Topic",
    values_to = "Value"
  ) %>%
  filter(Value == 1) %>%
  group_by(Topic) %>%
  summarise(n = n()) %>%
  mutate(share = n / sum(n))

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_paper_barplot.pdf")
ggplot(papers_long, aes(x = reorder(Topic, -n), y = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            vjust = -0.3, size = 5) +
  labs(x = NULL, y = "Number of papers") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 15),
    axis.text.y = element_text(size = 15),
    axis.title = element_text(size = 16))
dev.off()

## Different visualisation
install.packages("treemapify")
library(treemapify)

pdf("plots/Figure2_category_paper_tiles.pdf")
ggplot(papers_long, aes(area = share, fill = Topic, label = paste0(Topic, "\n", scales::percent(share, accuracy = 1)))) +
  geom_treemap(color = "white", size = 10) +
  geom_treemap_text(colour = "white", place = "centre", grow = F, reflow = TRUE) +
  scale_fill_manual(values = c(
  "Perceptions" = "#8B7B8B",
  "Behavior"    = "#EED8AE",
  "Priority"    = "#CDB5CD",
  "Policy"      = "#CD919E",
  "Health"      = "#8B475D",
  "None"        = "#F4A460")) +
  theme(legend.position = "none")
dev.off()

#-------------------------------------------------------------------------------
# surveys
#-------------------------------------------------------------------------------

# reshape to long format and count topics
surveys_long <- surveys %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health),
    names_to = "Topic",
    values_to = "Value"
  ) %>%
  filter(Value == 1) %>%
  group_by(Topic) %>%
  summarise(n = n()) %>%
  mutate(share = n / sum(n))

# bar plot with counts and percentage labels
pdf("plots/Figure2_category_survey_barplot.pdf")
ggplot(surveys_long, aes(x = reorder(Topic, -n), y = n, fill = Topic)) +
  geom_col(color = "black") +
  geom_text(aes(label = percent(share, accuracy = 1)),
            vjust = -0.3, size = 5) +
  labs(x = NULL, y = "Number of papers") +
  scale_fill_manual(values = c(
    "Perceptions" = "#8B7B8B",
    "Behavior"    = "#EED8AE",
    "Priority"    = "#CDB5CD",
    "Policy"      = "#CD919E",
    "Health"      = "#8B475D",
    "None"        = "#F4A460")) +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 15),
    axis.text.y = element_text(size = 15),
    axis.title = element_text(size = 16))
dev.off()

## Different approach
pdf("plots/Figure2_category_survey_tiles.pdf")
ggplot(surveys_long, aes(area = share, fill = Topic, label = paste0(Topic, "\n", scales::percent(share, accuracy = 1)))) +
  geom_treemap(color = "white", size = 10) +
  geom_treemap_text(colour = "white", place = "centre", grow = F, reflow = TRUE) +
  scale_fill_manual(values = c(
  "Perceptions" = "#8B7B8B",
  "Behavior"    = "#EED8AE",
  "Priority"    = "#CDB5CD",
  "Policy"      = "#CD919E",
  "Health"      = "#8B475D",
  "None"        = "#F4A460")) +
  theme(legend.position = "none")
dev.off()

```

## Visualizations by subgroup

```{r subgroupies}

#-------------------------------------------------------------------------------
# papers
#-------------------------------------------------------------------------------

install.packages("ggnewscale")
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggnewscale)

# 1. Prepare topic-level data (shares within income group)
topics_data <- papers %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(Income.group)) %>%
  group_by(Topic, Income.group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Income.group) %>%
  mutate(share = n / sum(n),
         total = sum(n)) %>%
  ungroup() %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))

# 2. Prepare totals row data (absolute counts)
totals_data <- topics_data %>%
  distinct(Income.group, total) %>%
  mutate(Topic = "Total papers")

# 3. Define consistent factor levels for the y-axis
topic_levels <- c("Total papers", unique(topics_data$Topic))

# 4. Plot with two independent fill scales
pdf("plots/Figure2_category_paper_by_income.pdf")
ggplot() +
  # --- top totals row ---
  geom_tile(data = totals_data,
            aes(x = Income.group, y = "Total papers", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = totals_data,
    aes(x = Income.group, y = "Total papers",
        label = paste0("n=", total),
        color = ifelse(total > max(topics_data$total) * 0.4, "white", "black")), size = 4) +
  scale_fill_gradient(
    name = "Total papers",
    low = "#EED8AE", high = "#8B864E",
    breaks = range(totals_data$total),         
  labels = range(totals_data$total)) +
  scale_color_identity() +

  ggnewscale::new_scale_fill() +  # reset fill scale

  # --- main topic rows ---
  geom_tile(data = topics_data,
            aes(x = Income.group, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data,
    aes(x = Income.group, y = Topic,
        label = scales::percent(share, accuracy = 1),
        color = ifelse(share > 0.25, "white", "black")),
    size = 4) +
  scale_fill_gradient(
    name = "Share within\nincome group",
    low = "#f7f4f9", high = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data$share), max(topics_data$share))) +
  scale_color_identity() +
labs(x = "Income group", y = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    legend.position = "top")

dev.off()

#-------------------------------------------------------------------------------
# surveys
#-------------------------------------------------------------------------------

# 1. Prepare topic-level data (shares within income group)
topics_data_s <- surveys %>%
  pivot_longer(
    cols = c(Perceptions, Behavior, Priority, Policy, Health),
    names_to = "Topic",
    values_to = "Value") %>%
  filter(Value == 1, !is.na(Income.group)) %>%
  group_by(Topic, Income.group) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(Income.group) %>%
  mutate(share = n / sum(n),
         total = sum(n)) %>%
  ungroup() %>%
  mutate(Income.group = factor(
    Income.group,
    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))

# 2. Prepare totals row data (absolute counts)
totals_data_s <- topics_data %>%
  distinct(Income.group, total) %>%
  mutate(Topic = "Total papers")

# 3. Define consistent factor levels for the y-axis
topic_levels_s <- c("Total items", unique(topics_data$Topic))

# 4. Plot with two independent fill scales
pdf("plots/Figure2_category_survey_by_income.pdf")
ggplot() +
  # --- top totals row ---
  geom_tile(data = totals_data_s,
            aes(x = Income.group, y = "Total items", fill = total),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = totals_data_s,
    aes(x = Income.group, y = "Total items",
        label = paste0("n=", total),
        color = ifelse(total > max(topics_data_s$total) * 0.4, "white", "black")), size = 4) +
  scale_fill_gradient(
    name = "Total items",
    low = "#EED8AE", high = "#8B864E",
    breaks = range(totals_data_s$total),         
  labels = range(totals_data_s$total)) +
  scale_color_identity() +

  ggnewscale::new_scale_fill() +  # reset fill scale

  # --- main topic rows ---
  geom_tile(data = topics_data_s,
            aes(x = Income.group, y = Topic, fill = share),
            color = "white", linewidth = 0.8) +
  geom_text(
    data = topics_data_s,
    aes(x = Income.group, y = Topic,
        label = scales::percent(share, accuracy = 1),
        color = ifelse(share > 0.25, "white", "black")),
    size = 4) +
  scale_fill_gradient(
    name = "Share within\nincome group",
    low = "#f7f4f9", high = "#8B475D",
    labels = scales::percent,
    breaks = c(min(topics_data_s$share), max(topics_data_s$share))) +
  scale_color_identity() +
labs(x = "Income group", y = NULL) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
    axis.text.y = element_text(size = 13),
    legend.position = "top")

dev.off()

```

### 2025-10-20 draftFiguretwo_Ella.Rmd
change width plots

@@
             vjust = -0.3, size = 5) +
-  labs(x = NULL, y = "Number of papers") +
+  labs(x = NULL, y = "Number of items") +
   scale_fill_manual(values = c(
@@
 # 4. Plot with two independent fill scales
-pdf("plots/Figure2_category_paper_by_income.pdf")
+pdf("plots/Figure2_category_paper_by_income.pdf", width = 10)
 ggplot() +
@@
 # 4. Plot with two independent fill scales
-pdf("plots/Figure2_category_survey_by_income.pdf")
+pdf("plots/Figure2_category_survey_by_income.pdf", width = 10)
 ggplot() +

### 2025-10-23 draftFiguresPaper.Rmd
fix dataset

@@
 papers_by_country <- datasetClassificationCountries_long_clean %>%
-  group_by(country, iso3, .drop = F) %>% 
+  group_by(iso3, .drop = F) %>% 
   count(iso3, name = "n_papers") %>%
@@
 library(rnaturalearth)
-
-# .
-# .
-# .
-
-write.csv(summaryDataset, "output/datasetPlots.csv", row.names = F)
+library(terra)
+
+world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))
+
+world_countries_df <- as.data.frame(world_countries)
+
+world_countries_df$iso_a3 <- ifelse(world_countries_df$iso_a3_eh == "FRA", "FRA",
+                                    ifelse(world_countries_df$iso_a3_eh == "NOR", "NOR", world_countries_df$iso_a3))
+
+summaryDataset2 <- merge(summaryDataset, world_countries_df %>% 
+                          select(continent, region_un, iso_a3), by.x = "iso3", by.y = "iso_a3", all.x = T, all.y = F)
+
+write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)
 

### 2025-10-23 draftFiguretwo_Ella.Rmd
color scale

@@
         label = paste0("n=", total),
-        color = ifelse(total > max(topics_data$total) * 0.4, "white", "black")), size = 4) +
+        color = ifelse(total > max(topics_data$total) * 0.4, "black", "black")), size = 4) +
   scale_fill_gradient(
     name = "Total papers",
-    low = "#EED8AE", high = "#8B864E",
+    low = "#F6DBA7", high = "#F7B32B",
     breaks = range(totals_data$total),         
@@
         label = paste0("n=", total),
-        color = ifelse(total > max(topics_data_s$total) * 0.4, "white", "black")), size = 4) +
+        color = ifelse(total > max(topics_data_s$total) * 0.4, "black", "black")), size = 4) +
   scale_fill_gradient(
     name = "Total items",
-    low = "#EED8AE", high = "#8B864E",
+    low = "#F6DBA7", high = "#F7B32B",
     breaks = range(totals_data_s$total),         

### 2025-10-23 draftFigure1.Rmd
Fig1 Fig2 v1

---
title: "draftFigure1"
output: html_document
---

```{r panels}

library(tidyverse)

dataset <- read.csv("output/datasetPlots.csv")

dataset$isDiff <- ifelse(dataset$isIndia == 1, "India",
                         ifelse(dataset$isChina == 1, "China", "Rest"))

summaryPaperInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_papers, na.rm = T)) %>% 
  filter(!is.na(Income.group))

summaryPaperInc$Income.group <- factor(
  summaryPaperInc$Income.group,
  levels = rev(c(
    "High income",
    "Upper middle income",
    "Lower middle income",
    "Low income"
  )),
  ordered = TRUE
)

# simple version of the plot
# summaryPaperInc %>% ggplot(aes(x = Income.group, y = nTotal))+
#   geom_bar(stat = "identity", aes(fill = isDiff))+
#   scale_fill_manual(values = c("#C19EAB", "#C19EAB", "#8B475D"))+
#   theme_minimal()+
#   labs(y = "Total number of papers", x = "Income group")+
#   theme(legend.position = "none")

totals <- summaryPaperInc %>%
  group_by(Income.group) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p1 <- summaryPaperInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#C19EAB", "#C19EAB", "#8B475D")) +
  theme_minimal() +
  labs(y = "Total number of papers", x = "Income group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p1

pdf("plots/Figure1_papers_by_income.pdf", width = 10)
p1
dev.off()

## then surveys
summarySurveysInc <- dataset %>% 
  group_by(Income.group, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(Income.group))

summarySurveysInc$Income.group <- factor(
  summarySurveysInc$Income.group,
  levels = rev(c(
    "High income",
    "Upper middle income",
    "Lower middle income",
    "Low income"
  )),
  ordered = TRUE
)

totals <- summarySurveysInc %>%
  group_by(Income.group) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p2 <- summarySurveysInc %>%
  ggplot(aes(x = Income.group, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = Income.group, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#E6D0F4", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "income Group") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p2

pdf("plots/Figure1_surveys_by_income.pdf", width = 10)
p2
dev.off()

## then papers per exposure

summaryPaperExp <- dataset %>% 
  group_by(categoryPollution, isDiff) %>% 
  summarize(nTotal = sum(n_papers, na.rm = T)) %>% 
  filter(!is.na(categoryPollution))

summaryPaperExp$categoryPollution <- ifelse(
  !is.na(summaryPaperExp$categoryPollution),
  paste0(toupper(substr(summaryPaperExp$categoryPollution, 1, 1)),
         substr(summaryPaperExp$categoryPollution, 2, nchar(summaryPaperExp$categoryPollution))),
  NA
)

summaryPaperExp$categoryPollution <- factor(
  summaryPaperExp$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

totals <- summaryPaperExp %>%
  group_by(categoryPollution) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p3 <- summaryPaperExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#C19EAB", "#A67384", "#8B475D")) +
  theme_minimal() +
  labs(y = "Total number of papers", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p3

pdf("plots/Figure1_papers_by_exposure.pdf", width = 10)
p3
dev.off()

## then surveys per exposure

summarySurveysExp <- dataset %>% 
  group_by(categoryPollution, isDiff) %>% 
  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
  filter(!is.na(categoryPollution))

summarySurveysExp$categoryPollution <- ifelse(
  !is.na(summarySurveysExp$categoryPollution),
  paste0(toupper(substr(summarySurveysExp$categoryPollution, 1, 1)),
         substr(summarySurveysExp$categoryPollution, 2, nchar(summarySurveysExp$categoryPollution))),
  NA
)

summarySurveysExp$categoryPollution <- factor(
  summarySurveysExp$categoryPollution,
  levels = rev(c(
    "Very high",
    "High",
    "Medium",
    "Low",
    "Very low"
  )),
  ordered = TRUE
)

totals <- summarySurveysExp %>%
  group_by(categoryPollution) %>%
  summarise(total = sum(nTotal), .groups = "drop")

p4 <- summarySurveysExp %>%
  ggplot(aes(x = categoryPollution, y = nTotal)) +
  geom_bar(stat = "identity", aes(fill = isDiff)) +
  geom_text(
    data = totals,
    aes(x = categoryPollution, y = total, label = total),
    vjust = -0.3, fontface = "bold"
  ) +
  scale_fill_manual(values = c("#E6D0F4", "#D4ABEF", "#B873E7")) +
  theme_minimal() +
  labs(y = "Total number of survey items", x = "Pollution exposure") +
  theme(legend.position = "none") +
  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping

p4

pdf("plots/Figure1_surveys_by_exposure.pdf", width = 10)
p4
dev.off()

```

```{r maps}

library(tidyverse)
library(rnaturalearth)
library(terra)
library(tidyterra)
library(scales)          # for label formatting

dataset <- read.csv("output/datasetPlots.csv")

world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))

world_countries$iso_a3 <- ifelse(world_countries$iso_a3_eh == "FRA", "FRA",
                                    ifelse(world_countries$iso_a3_eh == "NOR", "NOR", world_countries$iso_a3))

world_countries_merged <- merge(world_countries, dataset
                                , by.x = "iso_a3", by.y = "iso3", all.x = T, all.y = F)

world_countries_merged$n_papers <- ifelse(is.na(world_countries_merged$n_papers), 0, world_countries_merged$n_papers)

my_breaks = c(1, 5, 20, 50, 150)

world_no_antarctica <- 
  world_countries_merged %>%
  filter(name != "Antarctica")

# 2) Build the map with integer labels and no decimals:
map_plot_clean <- ggplot() +
  geom_spatvector(
    data = world_no_antarctica,
    aes(fill = n_papers),
    color = "grey80",
    size  = 0.1
  ) +
  scale_fill_gradientn(
    colours = c("#DCC9D2", "#8B475D", "#4F2A30"),
    name    = "Papers",
    trans   = "log10",
    breaks  = my_breaks,
    labels  = label_number(
      accuracy = 1,
      big.mark = ",",
      trim     = TRUE
    ),
    na.value = "grey90"
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title    = "Global Publication Counts by Country",
    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5)
  )

# Print it
map_plot_clean

ggsave(plot = map_plot_clean, filename = paste("plots/mapPapersCount.png", sep = ""),
       dpi=600, width = 16, height = 12, units='cm')

pdf("plots/Figure1_mapPapersCount.pdf", width = 10)
map_plot_clean
dev.off()

## then for surveys
world_countries_merged$n_surveys <- ifelse(is.na(world_countries_merged$n_surveys), 0, world_countries_merged$n_surveys)

my_breaks = c(1, 5, 10, 20)

# 1) Filter out Antarctica from your data:
world_no_antarctica <- 
  world_countries_merged %>%
  filter(name != "Antarctica")

# 2) Build the map with integer labels and no decimals:
map_plot_clean_survey <- ggplot() +
  geom_spatvector(
    data = world_no_antarctica,
    aes(fill = n_surveys),
    color = "grey70",
    size  = 0.1
  ) +
  scale_fill_gradientn(
    colours = c("#F3EBF8", "#D4ABEF", "#541F78"),
    name    = "Papers",
    trans   = "log10",
    breaks  = my_breaks,
    labels  = label_number(
      accuracy = 1,
      big.mark = ",",
      trim     = TRUE
    ),
    na.value = "grey85"
  ) +
  coord_sf(expand = FALSE) +
  labs(
    title    = "Global Survey Question Counts by Country",
    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
  ) +
  theme_void(base_size = 14) +
  theme(
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm"),
    plot.title       = element_text(face = "bold", hjust = 0.5),
    plot.subtitle    = element_text(hjust = 0.5)
  )

# Print it
map_plot_clean_survey

ggsave(plot = map_plot_clean_survey, filename = paste("plots/mapSurveyCount.png", sep = ""),
       dpi=600, width = 16, height = 12, units='cm')

pdf("plots/Figure1_mapSurveyCount.pdf", width = 10)
map_plot_clean_survey
dev.off()

```

### 2025-10-23 draftFiguresPaper.Rmd
Fig1 Fig2 v1

@@
     aes(fill = n_papers),
-    color = "grey70",
+    color = "grey80",
     size  = 0.1
   ) +
-  scale_fill_viridis_c(
+  scale_fill_gradientn(
+    colours = c("#DCC9D2", "#8B475D", "#4F2A30"),
     name    = "Papers",
-    # option  = "plasma",
     trans   = "log10",
     breaks  = my_breaks,
-    labels  = label_number(accuracy = 1,    # round to whole numbers
+    labels  = label_number(
+      accuracy = 1,
       big.mark = ",",
-                           trim     = TRUE),
-    na.value = "grey80"
+      trim     = TRUE
+    ),
+    na.value = "grey90"
   ) +
@@
 
-world_countries$iso_a3 <- ifelse(world_countries$sovereignt == "Somaliland", "SOM", world_countries$iso_a3)
-
-world_countries_merged <- merge(world_countries, surveys_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)
+# world_countries_merged <- merge(world_countries, surveys_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)
+
+world_countries$iso_a3 <- ifelse(world_countries$iso_a3_eh == "FRA", "FRA",
+                                    ifelse(world_countries$iso_a3_eh == "NOR", "NOR", world_countries$iso_a3))
+
+world_countries_merged <- merge(world_countries, surveys_by_country
+                                , by.x = "iso_a3", by.y = "iso3", all.x = T, all.y = F)
 

### 2025-10-24 draftFiguresPaper.Rmd
cleaning-up dataset and adding TWN HKG

@@
 
-### add income groups to the dataset
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
 # 2) get total papers by country
@@
   count(iso3, name = "n_papers") %>%
-  arrange(desc(n_papers)) %>% 
-  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
+  arrange(desc(n_papers))
 
@@
 datasetPollution_avg <- datasetPollution %>% 
-  filter(Year == 2019) %>% 
+  filter(Year %in% 2019) %>%
   group_by(Region) %>% 
   summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
+
+datasetPollution2 <- read.csv("dataNew/GlobalPM25-V5GL0502-Annual-1998-2023-wThresFrac.csv")
+
+datasetPollution2_avg <- datasetPollution2 %>% 
+  filter(Year %in% 2019) %>%
+  filter(Region %in% c("Hong Kong", "Taiwan")) %>% 
+  group_by(Region) %>% 
+  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
+
+datasetPollution_avg <- rbind(datasetPollution_avg, datasetPollution2_avg)
 
@@
 
-iso3_un_members <- c(
-  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
-  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
-  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
-  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
-  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
-  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
-  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
-  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
-  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
-  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
-  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
-  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
-  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
-  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
-  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
-  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
-  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
-  "VUT","VEN","VNM","YEM","ZMB","ZWE"
-)
-
-papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
-  filter(!is.na(iso3))
+# iso3_un_members <- c(
+#   "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
+#   "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
+#   "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
+#   "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
+#   "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
+#   "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
+#   "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
+#   "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
+#   "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
+#   "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
+#   "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
+#   "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
+#   "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
+#   "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
+#   "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
+#   "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
+#   "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
+#   "VUT","VEN","VNM","YEM","ZMB","ZWE"
+# )
+
+# papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
+#   filter(!is.na(iso3))
 
@@
 
-papers_by_country_pollution_filteredUN_survey <- merge(papers_by_country_pollution_filteredUN, surveys_by_country, by = "iso3", all.x = T, all.y = T)
-
-papers_by_country_pollution_filteredUN_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_filteredUN_survey$n_surveys), 0, papers_by_country_pollution_filteredUN_survey$n_surveys)
+papers_by_country_pollution_survey <- merge(papers_by_country_pollution, surveys_by_country, by = "iso3", all.x = T, all.y = T)
+
+papers_by_country_pollution_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_survey$n_surveys), 0, papers_by_country_pollution_survey$n_surveys)
 
@@
 
-summaryDataset <- merge(papers_by_country_pollution_filteredUN_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)
-
-summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-                  labels = c("very low", "low", "medium", "high", "very high"), 
+summaryDataset <- merge(papers_by_country_pollution_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)
+
+summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 5, 10, 15, 25, 35, 100), 
+                  labels = c("no pol", "very low", "low", "medium", "high", "very high"), 
                   right = FALSE)
+
+summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
+summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)
+
+### add income groups to the dataset
+
+incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
+incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
+
+summaryDataset <- merge(summaryDataset, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
 
@@
 
-summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
-summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)
-
-# add region from natural earth
+## add region from natural earth
 

### 2025-10-24 draftFiguresPaper.Rmd
categories pol

@@
 
-summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 5, 10, 15, 25, 35, 100), 
-                  labels = c("no pol", "very low", "low", "medium", "high", "very high"), 
+summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+                  labels = c("very low", "low", "medium", "high", "very high"), 
                   right = FALSE)

### 2025-10-27 draftFiguresPaper.Rmd
figure 2, by exposure

@@
 
+### get dataset figure 2 papers
+data_fig2_papers <- left_join(datasetClassificationCountries_long_clean, datasetPollution_ctr, by = "iso3")
+write.csv(data_fig2_papers,"dataNew/data_fig2_papers.csv", row.names = FALSE)
+
+### papers by country plus pollution
 papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)
@@
 
-## save this dataset for further analyses
-write.csv(surveyCountries_long_clean, "output/datasetSurveyCountries_long_clean.csv", row.names = F)
+### get dataset figure 2 suveys
+data_fig2_surveys <- left_join(surveyCountries_long_clean, datasetPollution_ctr, by = "iso3")
+write.csv(data_fig2_surveys,"dataNew/data_fig2_surveys.csv", row.names = FALSE)
 

### 2025-10-27 draftFiguretwo_Ella.Rmd
figure 2, by exposure

@@
 #-------------------------------------------------------------------------------
-papers <- read.csv2("output/datasetClassificationCountries_long_clean.csv", 
+papers <- read.csv2("dataNew/data_fig2_papers.csv", 
                     header = T, sep = ",", dec = ".")
 
-surveys <- read.csv2("output/datasetSurveyCountries_long_clean.csv", 
+surveys <- read.csv2("dataNew/data_fig2_surveys.csv", 
                      header = T, sep = ",", dec = ".")
@@
 papers <- papers[,c("ID", "country_clean",
-                    "Income.group", "Region", 
-                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "None")]
+                    "Income.group", "Region.x", 
+                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "None", "pm25Avg")]
 
-surveys <- surveys[,c("id", "country_clean", "year", "Income.group", "Region", 
-                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "Inclusion")]
+surveys <- surveys[,c("id", "country_clean", "year", "Income.group", "Region.x", 
+                      "Perceptions", "Behavior", "Priority", "Policy", "Health", "Inclusion", "pm25Avg")]
+
+## surveys: drop items for which Inclusion == 0
+surveys <- surveys[which(surveys$Inclusion == 1),]
+
+## subset survey without Flashbarometer
+surveys_sub <- surveys[which(surveys$source != "Flash Eurobarometer 360: Attitudes of Europeans towards air quality"),]
 
@@
 
-## Visualizations by subgroup
+## Visualizations by income group
 
-```{r subgroupies}
+```{r income group}
 
@@
   group_by(Income.group) %>%
-  mutate(share = n / sum(n),
-         total = sum(n)) %>%
+  mutate(share = n / sum(n)) %>%
   ungroup() %>%
@@
 # 2. Prepare totals row data (absolute counts)
-totals_data <- topics_data %>%
-  distinct(Income.group, total) %>%
-  mutate(Topic = "Total papers")
+totals_data <- papers %>%
+  filter(!is.na(Income.group)) %>%
+  count(Income.group, name = "total")
 
@@
 # 2. Prepare totals row data (absolute counts)
-totals_data_s <- topics_data %>%
-  distinct(Income.group, total) %>%
-  mutate(Topic = "Total papers")
+totals_data_s <- surveys %>%
+  filter(!is.na(Income.group)) %>%
+  count(Income.group, name = "total_items")
 
 # 3. Define consistent factor levels for the y-axis
-topic_levels_s <- c("Total items", unique(topics_data$Topic))
+topic_levels_s <- c("Total items", unique(topics_data_s$Topic))
 
@@
   geom_tile(data = totals_data_s,
-            aes(x = Income.group, y = "Total items", fill = total),
+            aes(x = Income.group, y = "Total items", fill = total_items),
             color = "white", linewidth = 0.8) +
@@
     aes(x = Income.group, y = "Total items",
-        label = paste0("n=", total),
-        color = ifelse(total > max(topics_data_s$total) * 0.4, "black", "black")), size = 4) +
+        label = paste0("n=", total_items),
+        color = ifelse(total_items > max(totals_data_s$total_items) * 0.4, "black", "black")), size = 4) +
   scale_fill_gradient(
@@
     low = "#F6DBA7", high = "#F7B32B",
-    breaks = range(totals_data_s$total),         
-  labels = range(totals_data_s$total)) +
+    breaks = range(totals_data_s$total_items),         
+  labels = range(totals_data_s$total_items)) +
   scale_color_identity() +
@@
 ```
+
+## Visualizations by exposure group
+
+```{r exposure group}
+
+#-------------------------------------------------------------------------------
+# get exposure groups for papers and surveys
+#-------------------------------------------------------------------------------
+papers$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
+                                right = FALSE)
+
+surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
+                                right = FALSE)
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within exposure group)
+topics_data_poll <- papers %>%
+  pivot_longer(
+    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
+    names_to = "Topic",
+    values_to = "Value") %>%
+  filter(Value == 1, !is.na(categoryPollution)) %>%
+  group_by(Topic, categoryPollution) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(categoryPollution) %>%
+  mutate(share = n / sum(n)) %>%
+  ungroup() %>%
+  mutate(categoryPollution = factor(
+    categoryPollution,
+    levels = c("Very low", "Low", "Medium", "High", "Very high")))
+
+# 2. Prepare totals row data (absolute counts)
+totals_data_poll <- papers %>%
+  filter(!is.na(categoryPollution)) %>%
+  count(categoryPollution, name = "total")
+
+# 3. Define consistent factor levels for the y-axis
+topic_levels <- c("Total papers", unique(topics_data$Topic))
+
+# 4. Plot with two independent fill scales
+pdf("plots/Figure2_category_paper_by_exposure.pdf", width = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = totals_data_poll,
+            aes(x = categoryPollution, y = "Total papers", fill = total),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = totals_data_poll,
+    aes(x = categoryPollution, y = "Total papers",
+        label = paste0("n=", total),
+        color = ifelse(total > max(topics_data_poll$total) * 0.4, "black", "black")), size = 4) +
+  scale_fill_gradient(
+    name = "Total papers",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(totals_data_poll$total),         
+  labels = range(totals_data_poll$total)) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = topics_data_poll,
+            aes(x = categoryPollution, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = topics_data_poll,
+    aes(x = categoryPollution, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 4) +
+  scale_fill_gradient(
+    name = "Share within\nexposure group",
+    low = "#f7f4f9", high = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data_poll$share), max(topics_data_poll$share))) +
+  scale_color_identity() +
+labs(x = "Exposure group", y = NULL) +
+  theme_minimal(base_size = 14) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
+    axis.text.y = element_text(size = 13),
+    legend.position = "top")
+
+dev.off()
+
+#-------------------------------------------------------------------------------
+# surveys
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within exposure group)
+topics_data_poll_s <- surveys %>%
+  pivot_longer(
+    cols = c(Perceptions, Behavior, Priority, Policy, Health),
+    names_to = "Topic",
+    values_to = "Value") %>%
+  filter(Value == 1, !is.na(categoryPollution)) %>%
+  group_by(Topic, categoryPollution) %>%
+  summarise(n = n(), .groups = "drop") %>%
+  group_by(categoryPollution) %>%
+  mutate(share = n / sum(n)) %>%
+  ungroup() %>%
+  mutate(categoryPollution = factor(
+    categoryPollution,
+    levels = c("Very low", "Low", "Medium", "High", "Very high")))
+
+# 2. Prepare totals row data (absolute counts)
+totals_data_poll_s <- surveys %>%
+  filter(!is.na(categoryPollution)) %>%
+  count(categoryPollution, name = "total_items")
+
+# 3. Define consistent factor levels for the y-axis
+topic_levels_s <- c("Total items", unique(topics_data_s$Topic))
+
+# 4. Plot with two independent fill scales
+pdf("plots/Figure2_category_survey_by_exposure.pdf", width = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = totals_data_poll_s,
+            aes(x = categoryPollution, y = "Total items", fill = total_items),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = totals_data_poll_s,
+    aes(x = categoryPollution, y = "Total items",
+        label = paste0("n=", total_items),
+        color = ifelse(total_items > max(totals_data_poll_s$total_items) * 0.4, "black", "black")), size = 4) +
+  scale_fill_gradient(
+    name = "Total items",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(totals_data_poll_s$total_items),         
+  labels = range(totals_data_poll_s$total_items)) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = topics_data_poll_s,
+            aes(x = categoryPollution, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = topics_data_poll_s,
+    aes(x = categoryPollution, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 4) +
+  scale_fill_gradient(
+    name = "Share within\nexposure group",
+    low = "#f7f4f9", high = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data_poll_s$share), max(topics_data_poll_s$share))) +
+  scale_color_identity() +
+labs(x = "Exposure group", y = NULL) +
+  theme_minimal(base_size = 14) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 13),
+    axis.text.y = element_text(size = 13),
+    legend.position = "top")
+
+dev.off()
+
+```

### 2025-10-27 draftFiguretwo_Ella.Rmd
update color scale surveys

@@
     name = "Share within\nincome group",
-    low = "#f7f4f9", high = "#8B475D",
+    low = "#F3EBF8", high = "#541F78",
     labels = scales::percent,
@@
 dev.off()
+
+#"#F3EBF8", "#D4ABEF", "#541F78"
 

### 2025-10-27 draftFiguresPaper.Rmd
cleanUp data code Fig2

@@
 surveys_by_country <- surveyCountries_long_clean %>%
+  filter(Inclusion == 1) %>% 
   group_by(iso3, .drop = F) %>% 
@@
 write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)
+
+```
+
+```{r long classification dataset - paper}
+
+library(tidyverse)
+library(countrycode)
+
+"%ni%" = Negate("%in%")
+
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
+
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
+
+datasetCountries <- datasetCountries %>% 
+  dplyr::select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
+
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
+                                        by = "ID") %>% 
+  dplyr::select(-reader)
+
+datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
+
+table(datasetClassificationCountries$country)
+
+# 1) split into one row per country
+datasetClassificationCountries_long <- datasetClassificationCountries %>%
+  separate_rows(country, sep = ",\\s*") %>%
+  mutate(countries = str_trim(country))
+
+datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
+                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
+                                                               datasetClassificationCountries_long$countries))
+# fix country names 
+# 1. Direct mapping
+datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+# add the year
+datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
+  dplyr::select(ID, Year)
+
+datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
+                                                   datasetYear,
+                                                   by = "ID", all.x = T, all.y = F)
+
+df_cat_long <- datasetClassificationCountries_long_clean %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  dplyr::select(-flag, -None)
+
+write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)
+
+```
+
+```{r long classification dataset - survey}
+
+library(tidyverse)
+library(countrycode)
+
+"%ni%" = Negate("%in%")
+
+surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
+                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
+                                                 surveyClassification$surveyedCountries)
+
+# 1) split into one row per country
+surveyCountries_long <- surveyClassification %>%
+  separate_rows(surveyedCountries, sep = ";\\s*") %>%
+  mutate(countries = str_trim(surveyedCountries))
+
+# fix country names first manual then with iso3
+
+surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
+                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
+                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
+                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
+                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
+                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
+                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
+                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
+
+# 1. Direct mapping
+surveyCountries_long_clean <- surveyCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+df_cat_long <- surveyCountries_long_clean %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  dplyr::select(-flag)
+
+write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)
 

### 2025-10-27 draftFiguretwo_Ella.Rmd
cleanUp data code Fig2

@@
 #-------------------------------------------------------------------------------
-papers <- read.csv2("dataNew/data_fig2_papers.csv", 
+papers <- read.csv2("output/paperClassifYear.csv", 
                     header = T, sep = ",", dec = ".")
 
-surveys <- read.csv2("dataNew/data_fig2_surveys.csv", 
+surveys <- read.csv2("output/surveyClassifYear.csv", 
                      header = T, sep = ",", dec = ".")
+
+dataSummary <- read.csv("output/datasetPlots.csv")
+
+dataSummary$categoryPollution <- ifelse(
+  !is.na(dataSummary$categoryPollution),
+  paste0(toupper(substr(dataSummary$categoryPollution, 1, 1)),
+         substr(dataSummary$categoryPollution, 2, nchar(dataSummary$categoryPollution))),
+  NA
+)
+
+dataSummary$categoryPollution <- factor(
+  dataSummary$categoryPollution,
+  levels = rev(c(
+    "Very high",
+    "High",
+    "Medium",
+    "Low",
+    "Very low"
+  )),
+  ordered = TRUE
+)
+
+papers <- merge(papers, dataSummary, by = "iso3", all.x = T, all.y = F)
+surveys <- merge(surveys, dataSummary, by = "iso3", all.x = T, all.y = F)
 
@@
 
-## Clean the data
+## Clean the data (would skip this?)
 
@@
 
+# install.packages("treemapify")
+library(tidyverse)
+library(scales)
+library(treemapify)
+
 #-------------------------------------------------------------------------------
@@
 
-# reshape to long format and count topics
+# count category
 papers_long <- papers %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
-    names_to = "Topic",
-    values_to = "Value"
-  ) %>%
-  filter(Value == 1) %>%
-  group_by(Topic) %>%
-  summarise(n = n()) %>%
-  mutate(share = n / sum(n))
+  count(category, name = "n") %>%   # counts per category
+  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
+  rename(Topic = category)
 
@@
 ## Different visualisation
-install.packages("treemapify")
-library(treemapify)
 
@@
 
-# reshape to long format and count topics
 surveys_long <- surveys %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health),
-    names_to = "Topic",
-    values_to = "Value"
-  ) %>%
-  filter(Value == 1) %>%
-  group_by(Topic) %>%
-  summarise(n = n()) %>%
-  mutate(share = n / sum(n))
+  count(category, name = "n") %>%   # counts per category
+  mutate(share = n / sum(n)) %>%    # sum(n) is the grand total here
+  rename(Topic = category)
 
@@
 
-install.packages("ggnewscale")
-library(dplyr)
-library(tidyr)
-library(ggplot2)
+# install.packages("ggnewscale")
+library(tidyverse)
+library(scales)
 library(ggnewscale)
@@
 topics_data <- papers %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
-    names_to = "Topic",
-    values_to = "Value") %>%
-  filter(Value == 1, !is.na(Income.group)) %>%
-  group_by(Topic, Income.group) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  group_by(Income.group) %>%
-  mutate(share = n / sum(n)) %>%
-  ungroup() %>%
+  count(Income.group, category, name = "n") %>%   # counts per category
+  rename(Topic = category) %>% 
+  filter(!is.na(Income.group)) %>% 
   mutate(Income.group = factor(
     Income.group,
-    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))
+    levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))) %>% 
+  group_by(Income.group) %>% 
+  mutate(share = n / sum(n))   # sum(n) is the grand total here
 
 # 2. Prepare totals row data (absolute counts)
-totals_data <- papers %>%
+totals_data <- dataSummary %>%
   filter(!is.na(Income.group)) %>%
-  count(Income.group, name = "total")
+  group_by(Income.group) %>% 
+  summarize(total = sum(n_papers, na.rm=T))
 
@@
 topics_data_s <- surveys %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health),
-    names_to = "Topic",
-    values_to = "Value") %>%
-  filter(Value == 1, !is.na(Income.group)) %>%
-  group_by(Topic, Income.group) %>%
-  summarise(n = n(), .groups = "drop") %>%
-  group_by(Income.group) %>%
-  mutate(share = n / sum(n)) %>%
-  ungroup() %>%
+  count(Income.group, category, name = "n") %>%   # counts per category
+  rename(Topic = category) %>% 
+  filter(!is.na(Income.group)) %>% 
   mutate(Income.group = factor(
     Income.group,
-    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")))
+    levels = c("Low income", "Lower middle income", "Upper middle income", "High income"))) %>% 
+  group_by(Income.group) %>% 
+  mutate(share = n / sum(n))   # sum(n) is the grand total here
 
 # 2. Prepare totals row data (absolute counts)
-totals_data_s <- surveys %>%
+totals_data_s <- dataSummary %>%
   filter(!is.na(Income.group)) %>%
-  count(Income.group, name = "total_items")
+  group_by(Income.group) %>% 
+  summarize(total_items = sum(n_surveys, na.rm=T))
 
@@
 #-------------------------------------------------------------------------------
-papers$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
-                                right = FALSE)
 
-surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-                                labels = c("Very low", "Low", "Medium", "High", "Very high"), 
-                                right = FALSE)
+## now is already in the papers dataset, could change values based on pollution value (pm25$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
+#                                 right = FALSE)
+
+# surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
+#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
+#                                 right = FALSE)
 
@@
 topics_data_poll <- papers %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health, None),
-    names_to = "Topic",
-    values_to = "Value") %>%
-  filter(Value == 1, !is.na(categoryPollution)) %>%
-  group_by(Topic, categoryPollution) %>%
-  summarise(n = n(), .groups = "drop") %>%
+  count(categoryPollution, category, name = "n") %>%   # counts per category
+  rename(Topic = category) %>% 
+  filter(!is.na(categoryPollution)) %>% 
   group_by(categoryPollution) %>% 
-  mutate(share = n / sum(n)) %>%
-  ungroup() %>%
-  mutate(categoryPollution = factor(
-    categoryPollution,
-    levels = c("Very low", "Low", "Medium", "High", "Very high")))
+  mutate(share = n / sum(n))   # sum(n) is the grand total here
 
 # 2. Prepare totals row data (absolute counts)
-totals_data_poll <- papers %>%
+totals_data_poll <- dataSummary %>%
   filter(!is.na(categoryPollution)) %>%
-  count(categoryPollution, name = "total")
+  group_by(categoryPollution) %>% 
+  summarize(total = sum(n_papers, na.rm=T))
 
 # 3. Define consistent factor levels for the y-axis
-topic_levels <- c("Total papers", unique(topics_data$Topic))
+topic_levels <- c("Total papers", unique(topics_data_poll$Topic))
 
@@
 topics_data_poll_s <- surveys %>%
-  pivot_longer(
-    cols = c(Perceptions, Behavior, Priority, Policy, Health),
-    names_to = "Topic",
-    values_to = "Value") %>%
-  filter(Value == 1, !is.na(categoryPollution)) %>%
-  group_by(Topic, categoryPollution) %>%
-  summarise(n = n(), .groups = "drop") %>%
+  count(categoryPollution, category, name = "n") %>%   # counts per category
+  rename(Topic = category) %>% 
+  filter(!is.na(categoryPollution)) %>% 
   group_by(categoryPollution) %>% 
-  mutate(share = n / sum(n)) %>%
-  ungroup() %>%
-  mutate(categoryPollution = factor(
-    categoryPollution,
-    levels = c("Very low", "Low", "Medium", "High", "Very high")))
+  mutate(share = n / sum(n))   # sum(n) is the grand total here
 
 # 2. Prepare totals row data (absolute counts)
-totals_data_poll_s <- surveys %>%
+totals_data_poll_s <- dataSummary %>%
   filter(!is.na(categoryPollution)) %>%
-  count(categoryPollution, name = "total_items")
+  group_by(categoryPollution) %>% 
+  summarize(total_items = sum(n_surveys, na.rm=T))
 
 # 3. Define consistent factor levels for the y-axis
-topic_levels_s <- c("Total items", unique(topics_data_s$Topic))
+topic_levels_s <- c("Total items", unique(topics_data_poll_s$Topic))
 

### 2025-10-27 draftFiguretwo_Ella.Rmd
minor changes Fig 2 (color scale surveys)

@@
     "data.table",
-    "scales")
+    "scales",
+    "tidyverse",
+    "treemapify",
+    "ggnewscale")
 
@@
 
-## Clean the data (would skip this?)
+## Filter (only Inclusion = 1 items)
 
@@
 #-------------------------------------------------------------------------------
-# consistent country names 
+# Inclusion = 1 for survey items 
 #-------------------------------------------------------------------------------
-
-## paper
-papers <- papers %>%
-  mutate(country_clean = recode(country,
-    # Turkey variants
-    "Turkey" = "Turkyie",
-    "Türkiye" = "Turkyie",
-
-    # UK variants
-    "UK" = "United Kingdom",
-    "England" = "United Kingdom",
-    "Wales" = "United Kingdom",
-
-    # Korea variants
-    "Korea" = "South Korea",
-
-    # Netherlands variants
-    "The Netherlands" = "Netherlands",
-
-    # Czechia variants
-    "Czech Republic" = "Czechia",
-
-    # US variants
-    "USA" = "United States",
-
-    .default = country  # keep as is otherwise
-  ))
-
-papers <- papers[which(papers$country_clean != "exclude"),]
-papers <- papers[which(papers$country_clean != "no info"),]
-
-## survey
-surveys <- surveys %>%
-  mutate(country_clean = recode(countries,
-    # Turkey variants
-    "Turkey" = "Turkyie",
-    "Türkiye" = "Turkyie",
-
-    # UK variants
-    "UK" = "United Kingdom",
-    "England" = "United Kingdom",
-    "Wales" = "United Kingdom",
-
-    # Korea variants
-    "Korea" = "South Korea",
-
-    # Netherlands variants
-    "The Netherlands" = "Netherlands",
-
-    # Czechia variants
-    "Czech Republic" = "Czechia",
-
-    # US variants
-    "USA" = "United States",
-
-    .default = countries  # keep as is otherwise
-  ))
-
-#-------------------------------------------------------------------------------
-# relevant columns
-#-------------------------------------------------------------------------------
-papers <- papers[,c("ID", "country_clean",
-                    "Income.group", "Region.x", 
-                    "Perceptions", "Behavior", "Priority", "Policy", "Health", "None", "pm25Avg")]
-
-surveys <- surveys[,c("id", "country_clean", "year", "Income.group", "Region.x", 
-                      "Perceptions", "Behavior", "Priority", "Policy", "Health", "Inclusion", "pm25Avg")]
 
@@
 ```{r visualise}
-
-# install.packages("treemapify")
-library(tidyverse)
-library(scales)
-library(treemapify)
 
@@
 #-------------------------------------------------------------------------------
-
 surveys_long <- surveys %>%
@@
 #-------------------------------------------------------------------------------
-
-# install.packages("ggnewscale")
-library(tidyverse)
-library(scales)
-library(ggnewscale)
 
@@
 
-#"#F3EBF8", "#D4ABEF", "#541F78"
-
 ```
@@
 ```{r exposure group}
-
-#-------------------------------------------------------------------------------
-# get exposure groups for papers and surveys
-#-------------------------------------------------------------------------------
-
-## now is already in the papers dataset, could change values based on pollution value (pm25$categoryPollution <- cut(papers$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
-#                                 right = FALSE)
-
-# surveys$categoryPollution <- cut(surveys$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-#                                 labels = c("Very low", "Low", "Medium", "High", "Very high"), 
-#                                 right = FALSE)
 
@@
     name = "Share within\nexposure group",
-    low = "#f7f4f9", high = "#8B475D",
+    low = "#F3EBF8", high = "#541F78",
     labels = scales::percent,

### 2025-12-05 draftFiguretwo_Ella.Rmd
figure2 shares

@@
 ---
+
+Redo figure 2 with the 'share of papers mentioning a topic' instead, so the percentages would go above 100% which is ok.
+Also for the other panels, calculate the percentages per subgroups that mention,
+which will also make each 'column' going above 100%
 
@@
 
+## Visualization counting the share of papers & survey items instead of the raw count for each category
+
+```{r}
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+papers_long <- papers %>%
+  group_by(category) %>%
+  summarise(n = n_distinct(ID)) %>%
+  rename(Topic = category) %>% 
+  mutate(share = n / length(unique(papers$ID)))
+
+# bar plot with counts and percentage labels
+pdf("plots/Figure2_category_paper_barplot_share.pdf", width = 12, height = 10)
+ggplot(papers_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
+  geom_col(color = "black") +
+  geom_text(aes(label = percent(share, accuracy = 1)),
+            hjust = -0.3, size = 9) +
+  labs(y = NULL, x = "Number of papers covering each category") +
+  scale_fill_manual(values = c(
+    "Perceptions" = "#8B7B8B",
+    "Behavior"    = "#EED8AE",
+    "Priority"    = "#CDB5CD",
+    "Policy"      = "#CD919E",
+    "Health"      = "#8B475D",
+    "None"        = "#F4A460")) +
+  theme_minimal() +
+  xlim(0, 450) +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(size = 22),
+    axis.text.x = element_text(size = 22),
+    axis.title = element_text(size = 23)
+    )
+dev.off()
+
+#-------------------------------------------------------------------------------
+# surveys
+#-------------------------------------------------------------------------------
+
+surveys_long <- surveys %>%
+  mutate(item_id = paste0(iso3, id)) %>% 
+  distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
+  count(category, name = "n") %>% 
+  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id)))) %>%
+  rename(Topic = category)
+
+# bar plot with counts and percentage labels
+pdf("plots/Figure2_category_survey_barplot_share.pdf", width = 12, height = 10)
+ggplot(surveys_long, aes(y = reorder(Topic, n), x = n, fill = Topic)) +
+  geom_col(color = "black") +
+  geom_text(aes(label = percent(share, accuracy = 1)),
+            hjust = -0.3, size = 9) +
+  labs(y = NULL, x = "Number of survey items covering each category") +
+  scale_fill_manual(values = c(
+    "Perceptions" = "#8B7B8B",
+    "Behavior"    = "#EED8AE",
+    "Priority"    = "#CDB5CD",
+    "Policy"      = "#CD919E",
+    "Health"      = "#8B475D",
+    "None"        = "#F4A460")) +
+  xlim(0, 1250) +
+  theme_minimal() +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(size = 22),
+    axis.text.x = element_text(size = 22),
+    axis.title = element_text(size = 23)
+    )
+dev.off()
+
+```
+
 ## Visualizations by income group
@@
 
+## Visualizations by income group, using share
+
 ## Visualizations by exposure group

### 2025-12-05 draftFiguretwo_Ella.Rmd
plots

@@
 
+```{r income share}
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within income group, based on unique papers)
+topics_data <- papers %>%
+  filter(!is.na(Income.group)) %>% 
+  distinct(ID, Income.group, category) %>%      # each paper counts once per (Income.group, category)
+  count(Income.group, category, name = "n_papers") %>%   # number of papers mentioning category
+  rename(Topic = category) %>%
+  mutate(Income.group = factor(
+    Income.group,
+    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+  ))
+
+# 2. Prepare totals row data (absolute counts of papers per income group)
+totals_data <- dataSummary %>%
+  filter(!is.na(Income.group)) %>%
+  group_by(Income.group) %>% 
+  summarize(total = sum(n_papers, na.rm = TRUE), .groups = "drop") %>%
+  mutate(Income.group = factor(
+    Income.group,
+    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+  ))
+
+topics_data <- topics_data %>%
+  left_join(totals_data, by = "Income.group") %>%
+  group_by(Income.group) %>%
+  mutate(
+    share = n_papers / total    # share of papers in that income group mentioning the topic
+  ) %>%
+  ungroup() %>%
+  mutate(Income.group = factor(
+    Income.group,
+    levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+  ),
+  Topic = factor(
+    Topic,
+    levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+  )
+  )
+
+# 3. Define consistent factor levels for the y-axis
+# topic_levels <- c("Total papers", unique(topics_data$Topic))
+
+pdf("plots/Figure2_category_paper_by_income_share.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = totals_data,
+            aes(x = Income.group, y = "Total papers", fill = total),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = totals_data,
+    aes(x = Income.group, y = "Total papers",
+        label = paste0("n=", total),
+        color = "black"),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Total papers",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(totals_data$total),
+    labels = range(totals_data$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = topics_data,
+            aes(x = Income.group, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = topics_data,
+    aes(x = Income.group, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        # label = paste0(
+        #     scales::percent(share, accuracy = 1), 
+        #     "\n(", n_papers, ")"
+        #   ),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Share of papers\nwithin income group ",
+    low = "#f7f4f9", high = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data$share), max(topics_data$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Income group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y = element_text(size = 22),
+    axis.title = element_text(size = 22),
+    legend.position = "top"
+  )
+dev.off()
+
+#-------------------------------------------------------------------------------
+# surveys
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within income group, based on unique survey items)
+surveys_topics <- surveys %>%
+  filter(!is.na(Income.group)) %>%
+  mutate(item_id = paste0(iso3, id)) %>%              # unique survey-level ID
+  distinct(item_id, Income.group, category) %>%       # each item counts once per (Income.group, category)
+  count(Income.group, category, name = "n_items") %>% # number of items mentioning category
+  rename(Topic = category) %>%
+  mutate(
+    Income.group = factor(
+      Income.group,
+      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+    ),
+    Topic = factor(
+      Topic,
+      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+    )
+  )
+
+# 2. Totals per income group (unique items)
+surveys_totals <- surveys %>%
+  filter(!is.na(Income.group)) %>%
+  mutate(item_id = paste0(iso3, id)) %>%
+  group_by(Income.group) %>%
+  summarise(
+    total = n_distinct(item_id),
+    .groups = "drop"
+  ) %>%
+  mutate(
+    Income.group = factor(
+      Income.group,
+      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+    )
+  )
+
+surveys_topics <- surveys_topics %>%
+  left_join(surveys_totals, by = "Income.group") %>%
+  group_by(Income.group) %>%
+  mutate(
+    share = n_items / total   # share of survey items in that income group mentioning the topic
+  ) %>%
+  ungroup()
+
+pdf("plots/Figure2_category_survey_by_income_share.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = surveys_totals,
+            aes(x = Income.group, y = "Total items", fill = total),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = surveys_totals,
+    aes(x = Income.group, y = "Total items",
+        label = paste0("n=", total),
+        color = "black"),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Total items",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(surveys_totals$total),
+    labels = range(surveys_totals$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = surveys_topics,
+            aes(x = Income.group, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = surveys_topics,
+    aes(x = Income.group, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        # label = paste0(
+        #     scales::percent(share, accuracy = 1), 
+        #     "\n(", n_papers, ")"
+        #   ),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Share of items\nwithin income group ",
+    low = "#f7f4f9", high = "#541F78",
+    labels = scales::percent,
+    breaks = c(min(surveys_topics$share), max(surveys_topics$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Income group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y = element_text(size = 22),
+    axis.title = element_text(size = 22),
+    legend.position = "top"
+  )
+dev.off()
+
+```
+
 ## Visualizations by exposure group
@@
 ```
+
+## per exposure level using shares
+
+```{r exposure level shares}
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within exposure group, based on unique papers)
+topics_data_poll <- papers %>%
+  filter(!is.na(categoryPollution)) %>%
+  distinct(ID, categoryPollution, category) %>%          # each paper counts once per (exposure, category)
+  count(categoryPollution, category, name = "n_papers") %>%   # number of papers mentioning category
+  rename(Topic = category) %>%
+  mutate(
+    # if you want a specific topic order, same as income plot:
+    Topic = factor(
+      Topic,
+      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+    )
+  )
+
+# 2. Prepare totals row data (absolute counts of papers per exposure group)
+totals_data_poll <- dataSummary %>%
+  filter(!is.na(categoryPollution)) %>%
+  group_by(categoryPollution) %>%
+  summarize(total = sum(n_papers, na.rm = TRUE), .groups = "drop")
+
+# 3. Add shares within exposure group (based on unique papers)
+topics_data_poll <- topics_data_poll %>%
+  left_join(totals_data_poll, by = "categoryPollution") %>%
+  group_by(categoryPollution) %>%
+  mutate(
+    share = n_papers / total   # share of papers in that exposure group mentioning the topic
+  ) %>%
+  ungroup()
+
+# 4. Plot with two independent fill scales
+pdf("plots/Figure2_category_paper_by_exposure_share.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(
+    data = totals_data_poll,
+    aes(x = categoryPollution, y = "Total papers", fill = total),
+    color = "white", linewidth = 0.8
+  ) +
+  geom_text(
+    data = totals_data_poll,
+    aes(
+      x = categoryPollution,
+      y = "Total papers",
+      label = paste0("n=", total),
+      color = "black"
+    ),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name  = "Total papers",
+    low   = "#F6DBA7",
+    high  = "#F7B32B",
+    breaks = range(totals_data_poll$total),
+    labels = range(totals_data_poll$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(
+    data = topics_data_poll,
+    aes(x = categoryPollution, y = Topic, fill = share),
+    color = "white", linewidth = 0.8
+  ) +
+  geom_text(
+    data = topics_data_poll,
+    aes(
+      x = categoryPollution,
+      y = Topic,
+      # OPTION 1: percentage only (like your current income plot)
+      label = scales::percent(share, accuracy = 1),
+      # OPTION 2: uncomment this instead for "percent\n(n_papers)"
+      # label = paste0(
+      #   scales::percent(share, accuracy = 1),
+      #   "\n(", n_papers, ")"
+      # ),
+      color = ifelse(share > 0.25, "white", "black")
+    ),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name   = "Share of papers\nwithin exposure group ",
+    low    = "#f7f4f9",
+    high   = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data_poll$share), max(topics_data_poll$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Exposure group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid   = element_blank(),
+    axis.text.x  = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y  = element_text(size = 22),
+    axis.title   = element_text(size = 22),
+    legend.position = "top"
+  )
+dev.off()
+
+#-------------------------------------------------------------------------------
+# surveys
+#-------------------------------------------------------------------------------
+
+# 1. Prepare topic-level data (shares within exposure group, based on unique survey items)
+topics_data_survey_poll <- surveys %>%
+  filter(!is.na(categoryPollution)) %>%
+  mutate(item_id = paste0(iso3, id)) %>%                      # unique survey item ID
+  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
+  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
+  rename(Topic = category) %>%
+  mutate(
+    # optional: topic order, same as your income-group plot
+    Topic = factor(
+      Topic,
+      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+    )
+    # optional: control exposure group order, e.g.:
+    # categoryPollution = factor(
+    #   categoryPollution,
+    #   levels = c("Air pollution", "Water pollution", "Other")  # adjust to your data
+    # )
+  )
+
+# 2. Prepare totals row data (absolute counts of unique items per exposure group)
+totals_data_survey_poll <- surveys %>%
+  filter(!is.na(categoryPollution)) %>%
+  mutate(item_id = paste0(iso3, id)) %>%
+  group_by(categoryPollution) %>%
+  summarise(
+    total = n_distinct(item_id),
+    .groups = "drop"
+  )
+  # optional: match factor ordering here too
+  # %>%
+  # mutate(
+  #   categoryPollution = factor(
+  #     categoryPollution,
+  #     levels = c("Air pollution", "Water pollution", "Other")
+  #   )
+  # )
+
+# 3. Add shares within exposure group (based on unique items)
+topics_data_survey_poll <- topics_data_survey_poll %>%
+  left_join(totals_data_survey_poll, by = "categoryPollution") %>%
+  group_by(categoryPollution) %>%
+  mutate(
+    share = n_items / total   # share of survey items in that exposure group mentioning the topic
+  ) %>%
+  ungroup()
+
+# 4. Plot with two independent fill scales
+pdf("plots/Figure2_category_item_by_exposure_share.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(
+    data = totals_data_survey_poll,
+    aes(x = categoryPollution, y = "Total items", fill = total),
+    color = "white", linewidth = 0.8
+  ) +
+  geom_text(
+    data = totals_data_survey_poll,
+    aes(
+      x = categoryPollution,
+      y = "Total items",
+      label = paste0("n=", total),
+      color = "black"
+    ),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name   = "Total items",
+    low    = "#F6DBA7",
+    high   = "#F7B32B",
+    breaks = range(totals_data_survey_poll$total),
+    labels = range(totals_data_survey_poll$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(
+    data = topics_data_survey_poll,
+    aes(x = categoryPollution, y = Topic, fill = share),
+    color = "white", linewidth = 0.8
+  ) +
+  geom_text(
+    data = topics_data_survey_poll,
+    aes(
+      x = categoryPollution,
+      y = Topic,
+      # OPTION 1: percentage only (mirroring your income-group code)
+      label = scales::percent(share, accuracy = 1),
+      # OPTION 2: percentage + underlying n_items
+      # label = paste0(
+      #   scales::percent(share, accuracy = 1),
+      #   "\n(", n_items, ")"
+      # ),
+      color = ifelse(share > 0.25, "white", "black")
+    ),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name   = "Share of items\nwithin exposure group ",
+    low    = "#f7f4f9",
+    high   = "#541F78",
+    labels = scales::percent,
+    breaks = c(min(topics_data_survey_poll$share), max(topics_data_survey_poll$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Exposure group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid    = element_blank(),
+    axis.text.x   = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y   = element_text(size = 22),
+    axis.title    = element_text(size = 22),
+    legend.position = "top"
+  )
+
+dev.off()
+
+```

### 2025-12-05 draftFiguresPaper.Rmd
fix bug year survey item

@@
 
+surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2017 - 2021", 2019, surveyCountries_long_clean$year)
+surveyCountries_long_clean$year <- ifelse(surveyCountries_long_clean$year == "2021 - 2022", 2021, surveyCountries_long_clean$year)
+
+surveyCountries_long_clean_t <- surveyCountries_long_clean %>%
+  # 1. build a single “years” string
+  mutate(
+    all_years = if_else(
+      is.na(repeatedMeasure) | repeatedMeasure == "",
+      as.character(year),
+      paste(year, repeatedMeasure, sep = ";")
+    ),
+    # drop every space character
+    all_years = str_remove_all(all_years, "\\s+")
+  ) %>%
+  separate_rows(all_years, sep = ";") %>%
+  mutate(year = as.integer(all_years)) # %>%
+  # dplyr::select(-all_years, -repeatedMeasure)
+
 # 2) get total surveys by country
-surveys_by_country <- surveyCountries_long_clean %>%
+surveys_by_country <- surveyCountries_long_clean_t %>%
   filter(Inclusion == 1) %>% 
@@
 
-df_cat_long <- surveyCountries_long_clean %>%
+surveyCountries_long_clean_yearly <- surveyCountries_long_clean
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)
+
+surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
+  # 1. build a single “years” string
+  mutate(
+    all_years = if_else(
+      is.na(repeatedMeasure) | repeatedMeasure == "",
+      as.character(year),
+      paste(year, repeatedMeasure, sep = ";")
+    ),
+    # drop every space character
+    all_years = str_remove_all(all_years, "\\s+")
+  ) %>%
+  separate_rows(all_years, sep = ";") %>%
+  mutate(year = as.integer(all_years)) %>%
+  dplyr::select(-all_years, -repeatedMeasure)
+
+df_cat_long <- surveyCountries_long_clean_yearly_t %>%
   pivot_longer(

### 2025-12-05 draftFiguretwo_Ella.Rmd
fix bug year survey item

@@
     axis.text.y = element_text(size = 15),
-    axis.title = element_text(size = 16))
+    axis.title = element_text(size = 16)
+    )
 dev.off()
@@
 surveys_long <- surveys %>%
-  mutate(item_id = paste0(iso3, id)) %>% 
+  mutate(item_id = paste0(iso3, id, year)) %>% 
   distinct(item_id, category) %>%     # collapse multiple rows for same ID–category pair
@@
     "None"        = "#F4A460")) +
-  xlim(0, 1250) +
+  xlim(0, 1850) +
   theme_minimal() +
@@
   filter(!is.na(Income.group)) %>%
-  mutate(item_id = paste0(iso3, id)) %>%              # unique survey-level ID
+  mutate(item_id = paste0(iso3, id, year)) %>%              # unique survey-level ID
   distinct(item_id, Income.group, category) %>%       # each item counts once per (Income.group, category)
@@
   filter(!is.na(Income.group)) %>%
-  mutate(item_id = paste0(iso3, id)) %>%
+  mutate(item_id = paste0(iso3, id, year)) %>%
   group_by(Income.group) %>%
@@
   filter(!is.na(categoryPollution)) %>%
-  mutate(item_id = paste0(iso3, id)) %>%                      # unique survey item ID
+  mutate(item_id = paste0(iso3, id, year)) %>%                      # unique survey item ID
   distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
@@
   filter(!is.na(categoryPollution)) %>%
-  mutate(item_id = paste0(iso3, id)) %>%
+  mutate(item_id = paste0(iso3, id, year)) %>%
   group_by(categoryPollution) %>%

### 2025-12-08 draftFiguresPaper.Rmd
reorganizing

@@
 
+```{r long classification dataset - paper}
+
+library(tidyverse)
+library(countrycode)
+
+"%ni%" = Negate("%in%")
+
+datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
+
+datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
+
+datasetCountries <- datasetCountries %>% 
+  dplyr::select(ID, countryPartialF) %>% 
+  rename(country = countryPartialF)
+
+datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
+                                        by = "ID") %>% 
+  dplyr::select(-reader)
+
+datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
+
+table(datasetClassificationCountries$country)
+
+# 1) split into one row per country
+datasetClassificationCountries_long <- datasetClassificationCountries %>%
+  separate_rows(country, sep = ",\\s*") %>%
+  mutate(countries = str_trim(country))
+
+datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
+                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
+                                                               datasetClassificationCountries_long$countries))
+# fix country names 
+# 1. Direct mapping
+datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+# add the year
+datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
+  dplyr::select(ID, Year)
+
+datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
+                                                   datasetYear,
+                                                   by = "ID", all.x = T, all.y = F)
+
+df_cat_long <- datasetClassificationCountries_long_clean %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  dplyr::select(-flag, -None)
+
+write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)
+
+```
+
+```{r long classification dataset - survey}
+
+library(tidyverse)
+library(countrycode)
+
+"%ni%" = Negate("%in%")
+
+surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
+                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
+                                                 surveyClassification$surveyedCountries)
+
+# 1) split into one row per country
+surveyCountries_long <- surveyClassification %>%
+  separate_rows(surveyedCountries, sep = ";\\s*") %>%
+  mutate(countries = str_trim(surveyedCountries))
+
+# fix country names first manual then with iso3
+
+surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
+                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
+                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
+                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
+                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
+                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
+                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
+                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
+
+# 1. Direct mapping
+surveyCountries_long_clean <- surveyCountries_long %>%
+  mutate(
+    iso3 = countrycode(countries,
+                       origin      = "country.name",
+                       destination = "iso3c")
+  ) %>% 
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+
+surveyCountries_long_clean_yearly <- surveyCountries_long_clean
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
+surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)
+
+surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
+  # 1. build a single “years” string
+  mutate(
+    all_years = if_else(
+      is.na(repeatedMeasure) | repeatedMeasure == "",
+      as.character(year),
+      paste(year, repeatedMeasure, sep = ";")
+    ),
+    # drop every space character
+    all_years = str_remove_all(all_years, "\\s+")
+  ) %>%
+  separate_rows(all_years, sep = ";") %>%
+  mutate(year = as.integer(all_years)) %>%
+  dplyr::select(-all_years, -repeatedMeasure)
+
+df_cat_long <- surveyCountries_long_clean_yearly_t %>%
+  pivot_longer(
+    cols = c(Perceptions, Priority, Policy, Health, Behavior),
+    names_to  = "category",
+    values_to = "flag"
+  ) %>%
+  # keep only “paper belongs to this category”
+  filter(flag == 1) %>%
+  dplyr::select(-flag)
+
+write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)
+
+```
+
 ```{r putting data together}
@@
 
+df_cat_long <- datasetClassificationCountries_long_clean %>%
+  # keep only “paper belongs to this category”
+  filter(sum(Perceptions, Priority, Policy, Health, Behavior) >= 1,
+         None != 1)
+
 # 2) get total papers by country
-papers_by_country <- datasetClassificationCountries_long_clean %>%
+papers_by_country <- df_cat_long %>%
   group_by(iso3, .drop = F) %>% 
@@
 write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)
-
-```
-
-```{r long classification dataset - paper}
-
-library(tidyverse)
-library(countrycode)
-
-"%ni%" = Negate("%in%")
-
-datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
-
-datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
-
-datasetCountries <- datasetCountries %>% 
-  dplyr::select(ID, countryPartialF) %>% 
-  rename(country = countryPartialF)
-
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
-                                        by = "ID") %>% 
-  dplyr::select(-reader)
-
-datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
-
-table(datasetClassificationCountries$country)
-
-# 1) split into one row per country
-datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ",\\s*") %>%
-  mutate(countries = str_trim(country))
-
-datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
-                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
-                                                               datasetClassificationCountries_long$countries))
-# fix country names 
-# 1. Direct mapping
-datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-# add the year
-datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
-  dplyr::select(ID, Year)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
-                                                   datasetYear,
-                                                   by = "ID", all.x = T, all.y = F)
-
-df_cat_long <- datasetClassificationCountries_long_clean %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  dplyr::select(-flag, -None)
-
-write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)
-
-```
-
-```{r long classification dataset - survey}
-
-library(tidyverse)
-library(countrycode)
-
-"%ni%" = Negate("%in%")
-
-surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
-
-surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
-                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
-                                                 surveyClassification$surveyedCountries)
-
-# 1) split into one row per country
-surveyCountries_long <- surveyClassification %>%
-  separate_rows(surveyedCountries, sep = ";\\s*") %>%
-  mutate(countries = str_trim(surveyedCountries))
-
-# fix country names first manual then with iso3
-
-surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
-                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
-                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
-                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
-                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
-                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
-                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
-                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
-
-# 1. Direct mapping
-surveyCountries_long_clean <- surveyCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-surveyCountries_long_clean_yearly <- surveyCountries_long_clean
-surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
-surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)
-
-surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
-  # 1. build a single “years” string
-  mutate(
-    all_years = if_else(
-      is.na(repeatedMeasure) | repeatedMeasure == "",
-      as.character(year),
-      paste(year, repeatedMeasure, sep = ";")
-    ),
-    # drop every space character
-    all_years = str_remove_all(all_years, "\\s+")
-  ) %>%
-  separate_rows(all_years, sep = ";") %>%
-  mutate(year = as.integer(all_years)) %>%
-  dplyr::select(-all_years, -repeatedMeasure)
-
-df_cat_long <- surveyCountries_long_clean_yearly_t %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  dplyr::select(-flag)
-
-write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)
 

### 2025-12-09 dataPrep.Rmd
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP

--- 2025-12-08 draftFiguresPaper.Rmd
@@
 ---
-title: "draftFiguresPaper"
-author: "Camille"
-date: "2025-05-23"
+title: "dataPrep"
 output: html_document
 ---
-
-```{r}
-library(tidyverse)
-"%ni%" = Negate("%in%")
-
-datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
-
-datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
-
-datasetCountries <- datasetCountries %>% 
-  select(ID, countryPartialF) %>% 
-  rename(country = countryPartialF)
-
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
-                                        by = "ID") %>% 
-  select(-reader)
-
-datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
-
-table(datasetClassificationCountries$country)
-
-# 1) split into one row per country
-datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ",\\s*") %>%
-  mutate(countries = str_trim(country))
-
-datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
-                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
-                                                               datasetClassificationCountries_long$countries))
-# fix country names 
-library(countrycode)
-
-# 1. Direct mapping
-datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-### add income groups to the dataset
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-# 2) get total papers by country
-papers_by_country <- datasetClassificationCountries_long_clean %>%
-  group_by(country, iso3, .drop = F) %>% 
-  count(iso3, name = "n_papers") %>%
-  arrange(desc(n_papers)) %>% 
-  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-papers_by_country$Income.group <- factor(
-  papers_by_country$Income.group,
-  levels = c(
-    "High income",
-    "Upper middle income",
-    "Lower middle income",
-    "Low income"
-  ),
-  ordered = TRUE
-)
-
-pCountryPaperIncome <- papers_by_country %>%
-  slice_max(n_papers, n = 20) %>%                     # top 30
-  mutate(
-    # 1) reorder by paper count (so bars go largest→smallest),
-    country = fct_reorder(country, n_papers)
-  ) %>%
-  ggplot(aes(x = country, y = n_papers, fill = Income.group)) +
-    geom_col() +
-    coord_flip() +
-    scale_fill_discrete(na.value = 'grey')+
-    labs(
-      x = "Country",
-      y = "Number of papers",
-      title = "Top 20 Countries by Paper Count"
-    ) +
-    theme_minimal()
-
-ggsave(plot = pCountryPaperIncome, filename = paste("plots/countryPapersIncome.png", sep = ""),
-       dpi=600, width = 16, height = 12, units='cm')
-
-#### make a map ####
-
-### getting the data together
-library(rnaturalearth)
-library(terra)
-library(tidyterra)
-
-world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))
-
-world_countries$iso_a3 <- ifelse(world_countries$sovereignt == "Somaliland", "SOM", world_countries$iso_a3)
-
-world_countries_merged <- merge(world_countries, papers_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)
-
-world_countries_merged$n_papers <- ifelse(is.na(world_countries_merged$n_papers), 0, world_countries_merged$n_papers)
-
-my_breaks = c(1, 5, 20, 50, 150)
-
-map_plot_reference <- ggplot() +
-  geom_spatvector(data = world_countries_merged, aes(fill = n_papers)) +
-  scale_fill_gradient(name = "nb of papers", trans = "log",
-                        breaks = my_breaks, labels = my_breaks)+
-  theme_void()
-
-plot(map_plot_reference)
-
-library(scales)          # for label formatting
-
-# 1) Filter out Antarctica from your data:
-world_no_antarctica <- 
-  world_countries_merged %>%
-  filter(name != "Antarctica")
-
-# 2) Build the map with integer labels and no decimals:
-map_plot_viridis_clean <- ggplot() +
-  geom_spatvector(
-    data = world_no_antarctica,
-    aes(fill = n_papers),
-    color = "grey80",
-    size  = 0.1
-  ) +
-  scale_fill_gradientn(
-    colours = c("#DCC9D2", "#8B475D", "#4F2A30"),
-    name    = "Papers",
-    trans   = "log10",
-    breaks  = my_breaks,
-    labels  = label_number(
-      accuracy = 1,
-      big.mark = ",",
-      trim     = TRUE
-    ),
-    na.value = "grey90"
-  ) +
-  coord_sf(expand = FALSE) +
-  labs(
-    title    = "Global Publication Counts by Country",
-    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
-  ) +
-  theme_void(base_size = 14) +
-  theme(
-    legend.position  = "bottom",
-    legend.key.width = unit(2, "cm"),
-    plot.title       = element_text(face = "bold", hjust = 0.5),
-    plot.subtitle    = element_text(hjust = 0.5)
-  )
-
-# Print it
-map_plot_viridis_clean
-
-ggsave(plot = map_plot_viridis_clean, filename = paste("plots/mapPapersCount.png", sep = ""),
-       dpi=600, width = 16, height = 12, units='cm')
-
-##### categories over time #####
-#
-
-# data with ID year 
-datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
-  select(ID, Year)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
-                                                   datasetYear,
-                                                   by = "ID", all.x = T, all.y = F)
-
-## save this dataset for further analyses
-write.csv(datasetClassificationCountries_long_clean, "output/datasetClassificationCountries_long_clean.csv", row.names = F)
-
-# 2) count papers per country-year
-yearly <- datasetClassificationCountries_long_clean %>%
-  count(Income.group, Year, name = "n_year")
-
-# 3) ensure every country has an entry for every year in your span
-years <- seq(min(datasetClassificationCountries_long_clean$Year),
-             max(datasetClassificationCountries_long_clean$Year))
-
-yearly_complete <- yearly %>%
-  complete(Income.group, Year = years, fill = list(n_year = 0))
-
-# 4) compute cumulative sum within each country
-cum <- yearly_complete %>%
-  group_by(Income.group) %>%
-  arrange(Year, .by_group = TRUE) %>%
-  mutate(n_cumulative = cumsum(n_year)) %>%
-  ungroup()
-
-cum$Income.group <- ifelse(is.na(cum$Income.group), "Unknown", cum$Income.group)
-
-cum$Income.group <- factor(
-  cum$Income.group,
-  levels = c(
-    "High income",
-    "Upper middle income",
-    "Lower middle income",
-    "Low income"
-    # "Unknown"
-  ),
-  ordered = TRUE
-)
-
-ppapersIncome <- subset(cum, !is.na(Income.group)) %>% 
-  ggplot(aes(x = Year, y = n_cumulative, col = Income.group)) +
-    geom_line(linewidth = 1) +
-    labs(
-      x = "Year",
-      y = "Cumulative # of Papers",
-      title = "Growth of Papers by income group"
-    ) +
-    theme_minimal()
-
-ggsave(plot = ppapersIncome, filename = paste("plots/incomePaperYear.png", sep = ""),
-       dpi=600, width = 16, height = 8, units='cm')
-
-# then cumulative for different categories of papers
-
-# 1) pivot your dummy columns into long form
-df_cat_long <- datasetClassificationCountries_long_clean %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  select(-flag)
-
-# 2) count per year × income_group × category
-yearly_cat <- df_cat_long %>%
-  count(Year, Income.group, category, name = "n_papers")
-
-# 3) fill zeros so lines don’t break
-all_years <- seq(min(datasetClassificationCountries_long_clean$Year), max(datasetClassificationCountries_long_clean$Year))
-
-yearly_cat <- yearly_cat %>%
-  complete(
-    Year, Income.group, category,
-    fill = list(n_papers = 0)
-  )
-
-yearly_cat$Income.group <- factor(
-  yearly_cat$Income.group,
-  levels = c(
-    "High income",
-    "Upper middle income",
-    "Lower middle income",
-    "Low income"
-    # "Unknown"
-  ),
-  ordered = TRUE
-)
-
-yearly_cat <- yearly_cat %>%
-  group_by(Income.group, category) %>%
-  arrange(Year, .by_group = TRUE) %>%
-  mutate(n_cum = cumsum(n_papers)) %>%
-  ungroup() %>% 
-  # reorder so levels go from lowest to highest n_cum at the final Year
-  mutate(category = fct_reorder2(category, Year, n_cum))
-
-# 4) plot: one facet per category, colored by income group
-ppapersIncomeYear <- ggplot(subset(yearly_cat, !is.na(Income.group)),
-       aes(x = Year, y = n_cum, color = category)) +
-  geom_line(linewidth = 1) +
-  facet_wrap(~ Income.group, scales = "free_y") +
-  theme_minimal()
-
-ggsave(plot = ppapersIncomeYear, filename = paste("plots/incomeCategoryPapersYear.png", sep = ""),
-       dpi=600, width = 20, height = 12, units='cm')
-
-# a) sum across incomes, then compute cumulative
-yearly_cat_agg <- yearly_cat %>%
-  group_by(Year, category) %>%
-  summarise(n_year = sum(n_papers), .groups="drop") %>%
-  arrange(category, Year) %>%
-  group_by(category) %>%
-  mutate(n_cum = cumsum(n_year)) %>%
-  ungroup()
-
-# b) plot one line per category
-ppapersYear <- ggplot(yearly_cat_agg,
-       aes(x = Year, y = n_cum, color = category)) +
-  geom_line(linewidth = 1) +
-  theme_minimal() +
-  labs(
-    x     = "Year",
-    y     = "Cumulative # of Papers",
-    color = "Category",
-    title = "Cumulative Papers by Category (all incomes combined)"
-  )
-
-ggsave(plot = ppapersYear, filename = paste("plots/countryPapersYear.png", sep = ""),
-       dpi=600, width = 16, height = 8, units='cm')
-
-```
 
@@
   ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
+  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia")) # virtually filtering for papers that would not have a matching country
 
@@
 surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
+
+surveyClassification <- surveyClassification %>% 
+  filter(Inclusion == 1)
 
@@
 
-# iso3_un_members <- c(
-#   "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
-#   "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
-#   "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
-#   "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
-#   "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
-#   "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
-#   "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
-#   "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
-#   "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
-#   "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
-#   "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
-#   "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
-#   "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
-#   "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
-#   "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
-#   "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
-#   "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
-#   "VUT","VEN","VNM","YEM","ZMB","ZWE"
-# )
-
-# papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
-#   filter(!is.na(iso3))
-
 ## adding survey nb
@@
 
+summaryDataset2 <- summaryDataset2 %>% 
+  filter(!(iso3 == "FRA" & continent == "Seven seas (open ocean)")) %>% 
+  filter(!(iso3 == "CYP" & countryName == "Northern Cyprus"))
+
 write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)
@@
 ```
-
-```{r linking nb of papers with pollution levels over space and time?}
-
-library(tidyverse)
-"%ni%" = Negate("%in%")
-
-datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
-
-datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
-
-datasetCountries <- datasetCountries %>% 
-  select(ID, countryPartialF) %>% 
-  rename(country = countryPartialF)
-
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
-                                        by = "ID") %>% 
-  select(-reader)
-
-datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
-
-table(datasetClassificationCountries$country)
-
-# 1) split into one row per country
-datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ",\\s*") %>%
-  mutate(countries = str_trim(country))
-
-datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
-                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
-                                                               datasetClassificationCountries_long$countries))
-# fix country names 
-library(countrycode)
-
-# 1. Direct mapping
-datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-### add income groups to the dataset
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-# 2) get total papers by country
-papers_by_country <- datasetClassificationCountries_long_clean %>%
-  group_by(country, iso3, .drop = F) %>% 
-  count(iso3, name = "n_papers") %>%
-  arrange(desc(n_papers)) %>% 
-  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")
-
-datasetPollution_avg <- datasetPollution %>% 
-  filter(Year %in% 2019) %>%
-  group_by(Region) %>% 
-  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
-
-datasetPollution2 <- read.csv("dataNew/GlobalPM25-V5GL0502-Annual-1998-2023-wThresFrac.csv")
-
-datasetPollution2_avg <- datasetPollution2 %>% 
-  filter(Year %in% 2019) %>%
-  filter(Region %in% c("Hong Kong", "Taiwan")) %>% 
-  group_by(Region) %>% 
-  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
-
-datasetPollution_avg <- rbind(datasetPollution_avg, datasetPollution2_avg)
-
-datasetPollution_ctr <- datasetPollution_avg %>%
-  mutate(
-    iso3 = countrycode(Region,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(!is.na(iso3))
-
-### get dataset figure 2 papers
-data_fig2_papers <- left_join(datasetClassificationCountries_long_clean, datasetPollution_ctr, by = "iso3")
-write.csv(data_fig2_papers,"dataNew/data_fig2_papers.csv", row.names = FALSE)
-
-### papers by country plus pollution
-papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)
-
-fullCountrylist <- unique(countrycode::codelist$iso3c)
-
-papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)
-
-papers_by_country_pollution_filtered <- subset(papers_by_country_pollution, iso3 %in% fullCountrylist) %>% 
-  filter(!is.na(iso3))
-
-papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()
-
-papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth()
-
-papers_by_country_pollution_filtered %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-subset(papers_by_country_pollution_filtered, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filtered %>% 
-  filter(n_papers > 0) %>%
-  filter(iso3 != "CHN") %>% 
-  ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filtered %>% 
-  # filter(n_papers > 0) %>%
-  filter(iso3 != "CHN") %>% 
-  ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-iso3_un_members <- c(
-  "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
-  "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
-  "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
-  "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
-  "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
-  "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
-  "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
-  "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
-  "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
-  "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
-  "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
-  "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
-  "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
-  "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
-  "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
-  "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
-  "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
-  "VUT","VEN","VNM","YEM","ZMB","ZWE"
-)
-
-papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
-  filter(!is.na(iso3))
-
-papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()
-
-papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth()
-
-papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filteredUN %>% ggplot(aes(x = pm25Avg, y = (n_papers)))+
-  geom_point()+
-  geom_smooth(method = "lm")+
-  ylim(0,40)
-
-subset(papers_by_country_pollution_filteredUN, n_papers > 0) %>% ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filteredUN %>% 
-  filter(n_papers > 0) %>%
-  filter(iso3 != "CHN") %>% 
-  ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filteredUN %>% 
-  # filter(n_papers > 0) %>%
-  filter(iso3 != "CHN") %>% 
-  ggplot(aes(x = pm25Avg, y = n_papers))+
-  geom_point()+
-  geom_smooth(method = "lm")
-
-papers_by_country_pollution_filteredUN %>% ggplot(aes(x = n_papers > 0, y = pm25Avg))+
-  geom_boxplot()
-
-papers_by_country_pollution_filteredUN$categoryPollution <- cut(papers_by_country_pollution_filteredUN$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-                  labels = c("very low", "low", "medium", "high", "very high"), 
-                  right = FALSE)
-
-papers_by_country_pollution_filteredUN %>% 
-  # filter(iso3 != "CHN") %>%
-  ggplot(aes(x = categoryPollution, y = log(n_papers)))+
-  geom_boxplot()
-
-papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
-  mutate(paperBin = ifelse(n_papers > 0, 1, 0)) %>% 
-  group_by(categoryPollution) %>% 
-  summarize(proportionPaper = sum(paperBin, na.rm=T)/n(),
-            nbCountries = n())
-
-papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = proportionPaper))+
-  geom_boxplot()+
-  ylim(0,1)
-
-papers_by_country_pollution_filteredUN_summary <- papers_by_country_pollution_filteredUN %>% 
-  group_by(categoryPollution) %>% 
-  summarize(papersPerCountry = sum(n_papers, na.rm=T)/n(),
-            nbCountries = n(),
-            medianPapersPerCountry = median(n_papers, na.rm=T))
-
-papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = medianPapersPerCountry))+
-  geom_boxplot()
-
-table(papers_by_country_pollution_filteredUN$categoryPollution)
-
-papers_by_country_pollution_filteredUN_summary %>% ggplot(aes(x = categoryPollution, y = papersPerCountry))+
-  geom_boxplot()
-
-TEST <- papers_by_country_pollution_filteredUN %>% 
-  filter(categoryPollution == "very high")
-
-# ylim(0,1)
-```
-
-```{r survey classification}
-
-library(tidyverse)
-"%ni%" = Negate("%in%")
-
-surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
-
-surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
-                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
-                                                 surveyClassification$surveyedCountries)
-
-# 1) split into one row per country
-surveyCountries_long <- surveyClassification %>%
-  separate_rows(surveyedCountries, sep = ";\\s*") %>%
-  mutate(countries = str_trim(surveyedCountries))
-
-# fix country names first manual then with iso3
-
-surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
-                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
-                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
-                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
-                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
-                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
-                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
-                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
-
-library(countrycode)
-
-# 1. Direct mapping
-surveyCountries_long_clean <- surveyCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-surveyCountries_long_clean <- merge(surveyCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-### get dataset figure 2 suveys
-data_fig2_surveys <- left_join(surveyCountries_long_clean, datasetPollution_ctr, by = "iso3")
-write.csv(data_fig2_surveys,"dataNew/data_fig2_surveys.csv", row.names = FALSE)
-
-# 2) get total surveys by country
-surveys_by_country <- surveyCountries_long_clean %>%
-  group_by(countries, iso3, .drop = F) %>% 
-  count(iso3, name = "n_surveys") %>%
-  arrange(desc(n_surveys)) %>% 
-  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-surveys_by_country$Income.group <- factor(
-  surveys_by_country$Income.group,
-  levels = c(
-    "High income",
-    "Upper middle income",
-    "Lower middle income",
-    "Low income"
-  ),
-  ordered = TRUE
-)
-
-psurveyCountry <- surveys_by_country %>%
-  slice_max(n_surveys, n = 40) %>%                     # top 30
-  mutate(
-    # 1) reorder by paper count (so bars go largest→smallest),
-    country = fct_reorder(countries, n_surveys),
-    # 2) then shove "Unknown" to the very end
-    # country = fct_relevel(country, "Unknown", after = 0)
-  ) %>%
-  ggplot(aes(x = country, y = n_surveys, fill = Income.group)) +
-    geom_col() +
-    coord_flip() +
-    scale_fill_discrete(na.value = 'grey')+
-    labs(
-      x = "Country",
-      y = "Number of Surveys",
-      title = "Top 40 Countries by Survey Question Count"
-    ) +
-    theme_minimal()
-
-ggsave(plot = psurveyCountry, filename = paste("plots/countrySurveyIncome.png", sep = ""),
-       dpi=600, width = 16, height = 12, units='cm')
-
-df_cat_long <- surveyCountries_long_clean %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  select(-flag)
-
-# 2) count per year × income_group × category
-overall_cat <- df_cat_long %>%
-  count(Income.group, category, name = "n_surveys")
-
-# 4) plot: one facet per category, colored by income group
-ggplot(overall_cat,
-       aes(y = n_surveys, x = category, fill = category)) +
-  geom_col() +
-  facet_wrap(~ Income.group, scales = "free_y") +
-  theme_minimal()
-
-## same but over time
-
-surveyCountries_long_clean_yearly <- surveyCountries_long_clean
-surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2018, surveyCountries_long_clean_yearly$year)
-surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2021 - 2022", 2021, surveyCountries_long_clean_yearly$year)
-
-surveyCountries_long_clean_yearly_t <- surveyCountries_long_clean_yearly %>%
-  # 1. build a single “years” string
-  mutate(
-    all_years = if_else(
-      is.na(repeatedMeasure) | repeatedMeasure == "",
-      as.character(year),
-      paste(year, repeatedMeasure, sep = ";")
-    ),
-    # drop every space character
-    all_years = str_remove_all(all_years, "\\s+")
-  ) %>%
-  separate_rows(all_years, sep = ";") %>%
-  mutate(year_num = as.integer(all_years)) %>%
-  select(-all_years, -repeatedMeasure)
-
-df_cat_long <- surveyCountries_long_clean_yearly_t %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  select(-flag)
-
-# 2) count per year × income_group × category
-yearly_cat <- df_cat_long %>%
-  count(year_num, Income.group, category, name = "n_surveys")
-
-# 3) fill zeros so lines don’t break
-all_years <- seq(min(surveyCountries_long_clean_yearly_t$year_num), max(surveyCountries_long_clean_yearly_t$year_num))
-
-yearly_cat_filled <- yearly_cat %>%
-  complete(
-    year_num = all_years,
-    Income.group, category,
-    fill = list(n_surveys = 0)
-  )
-
-yearly_cat_filled <- yearly_cat_filled %>%
-  group_by(Income.group, category) %>%
-  arrange(year_num, .by_group = TRUE) %>%
-  mutate(n_cum = cumsum(n_surveys)) %>%
-  ungroup() %>% 
-  # reorder so levels go from lowest to highest n_cum at the final Year
-  mutate(category = fct_reorder2(category, year_num, n_cum))
-
-# 4) plot: one facet per income group, colored by category
-ggplot(yearly_cat_filled,
-       aes(x = year_num, y = n_cum, color = category)) +
-  # geom_point()+
-  geom_line(linewidth = 2) +
-  facet_wrap(~ Income.group, scales = "free_y") +
-  theme_minimal()
-
-yearly_cat_filled$Income.group <- factor(
-  yearly_cat_filled$Income.group,
-  levels = c(
-    "High income",
-    "Upper middle income",
-    "Lower middle income",
-    "Low income"
-  ),
-  ordered = TRUE
-)
-
-yearly_cat_filled$category <- factor(
-  yearly_cat_filled$category,
-  levels = c(
-    "Perceptions",
-    "Health",
-    "Behavior",
-    "Policy",
-    "Priority"
-  ),
-  ordered = TRUE
-)
-
-pyear_survey <- ggplot(yearly_cat_filled,
-       aes(x = year_num, y = n_cum, color = category)) +
-  # geom_point()+
-  geom_line(linewidth = 1) +
-  facet_wrap(~ Income.group, scales = "free_y") +
-  theme_minimal()
-
-ggsave(plot = pyear_survey, filename = paste("plots/yearSurveyIncome.png", sep = ""),
-       dpi=600, width = 16, height = 8, units='cm')
-
-```
-
-```{r make a map for the survey data}
-
-#### make a map ####
-
-### getting the data together
-library(rnaturalearth)
-library(terra)
-library(tidyterra)
-library(tidyverse)
-
-"%ni%" = Negate("%in%")
-
-surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
-
-surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
-                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
-                                                 surveyClassification$surveyedCountries)
-
-# 1) split into one row per country
-surveyCountries_long <- surveyClassification %>%
-  separate_rows(surveyedCountries, sep = ";\\s*") %>%
-  mutate(countries = str_trim(surveyedCountries))
-
-# fix country names first manual then with iso3
-
-surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
-                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
-                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
-                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
-                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
-                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
-                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
-                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
-
-library(countrycode)
-
-# 1. Direct mapping
-surveyCountries_long_clean <- surveyCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-surveyCountries_long_clean <- merge(surveyCountries_long_clean, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-# 2) get total surveys by country
-surveys_by_country <- surveyCountries_long_clean %>%
-  group_by(countries, iso3, .drop = F) %>% 
-  count(iso3, name = "n_surveys") %>%
-  arrange(desc(n_surveys)) %>% 
-  merge(., incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))
-
-# world_countries_merged <- merge(world_countries, surveys_by_country, by.x = "adm0_a3", by.y = "iso3", all.x = T)
-
-world_countries$iso_a3 <- ifelse(world_countries$iso_a3_eh == "FRA", "FRA",
-                                    ifelse(world_countries$iso_a3_eh == "NOR", "NOR", world_countries$iso_a3))
-
-world_countries_merged <- merge(world_countries, surveys_by_country
-                                , by.x = "iso_a3", by.y = "iso3", all.x = T, all.y = F)
-
-world_countries_merged$n_surveys <- ifelse(is.na(world_countries_merged$n_surveys), 0, world_countries_merged$n_surveys)
-
-my_breaks = c(1, 5, 10, 20)
-
-map_plot_reference <- ggplot() +
-  geom_spatvector(data = world_countries_merged, aes(fill = n_surveys)) +
-  scale_fill_gradient(name = "nb of surveys", trans = "log",
-                        breaks = my_breaks, labels = my_breaks)+
-  theme_void()
-
-plot(map_plot_reference)
-
-library(scales)          # for label formatting
-
-# 1) Filter out Antarctica from your data:
-world_no_antarctica <- 
-  world_countries_merged %>%
-  filter(name != "Antarctica")
-
-# 2) Build the map with integer labels and no decimals:
-map_plot_viridis_clean_survey <- ggplot() +
-  geom_spatvector(
-    data = world_no_antarctica,
-    aes(fill = n_surveys),
-    color = "grey70",
-    size  = 0.1
-  ) +
-  scale_fill_viridis_c(
-    name    = "Survey Questions",
-    # option  = "plasma",
-    trans   = "log10",
-    breaks  = my_breaks,
-    labels  = label_number(accuracy = 1,    # round to whole numbers
-                           big.mark = ",",
-                           trim     = TRUE),
-    na.value = "grey80"
-  ) +
-  coord_sf(expand = FALSE) +
-  labs(
-    title    = "Global Survey Question Counts by Country",
-    # subtitle = "Log₁₀ scale, viridis “plasma” palette"
-  ) +
-  theme_void(base_size = 14) +
-  theme(
-    legend.position  = "bottom",
-    legend.key.width = unit(2, "cm"),
-    plot.title       = element_text(face = "bold", hjust = 0.5),
-    plot.subtitle    = element_text(hjust = 0.5)
-  )
-
-# Print it
-map_plot_viridis_clean_survey
-
-ggsave(plot = map_plot_viridis_clean_survey, filename = paste("plots/mapSurveyCount.png", sep = ""),
-       dpi=600, width = 16, height = 12, units='cm')
-
-```

### 2025-12-09 draftFigure2.Rmd
reorganizing
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP

--- 2025-12-05 draftFiguretwo_Ella.Rmd
@@
 ---
-
-Redo figure 2 with the 'share of papers mentioning a topic' instead, so the percentages would go above 100% which is ok.
-Also for the other panels, calculate the percentages per subgroups that mention,
-which will also make each 'column' going above 100%
 
@@
 
+papers <- papers %>% 
+  filter(!is.na(iso3))
+
+surveys <- surveys %>% 
+  filter(!is.na(iso3))
+
+dataSummary <- dataSummary %>% 
+  filter(!is.na(iso3))
+
 dataSummary$categoryPollution <- ifelse(
@@
 ## surveys: drop items for which Inclusion == 0
-surveys <- surveys[which(surveys$Inclusion == 1),]
+surveys <- surveys[which(surveys$Inclusion == 1),] # done in the data prep section
 
@@
   count(category, name = "n") %>% 
-  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id)))) %>%
+  mutate(share = n / length(unique(paste0(surveys$iso3, surveys$id, surveys$year)))) %>%
   rename(Topic = category)
@@
 
+# Visualizations by income group, with proportional weighting
+
+```{r visualisation income weighting}
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+## 1. Compute country-weights per paper x income group -----------------------
+
+# For each paper, how many distinct countries are there in total,
+# and how many of those are in each income group?
+paper_country_weights <- papers %>%
+  distinct(ID, country, Income.group) %>%  # ignore category here
+  group_by(ID) %>%
+  mutate(n_countries_total = n_distinct(country)) %>%
+  group_by(ID, Income.group, n_countries_total) %>%
+  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
+  mutate(country_weight = n_countries_group / n_countries_total)
+# Example: if a paper has 2 countries (1 low, 1 high) then
+# country_weight = 0.5 for low income and 0.5 for high income.
+
+## 2. Attach weights back to the (paper x income group x category) rows ------
+
+papers_weighted <- papers %>%
+  left_join(
+    paper_country_weights %>%
+      dplyr::select(ID, Income.group, country_weight),
+    by = c("ID", "Income.group")
+  )
+
+## 3. Compute weighted category shares within each income group --------------
+
+topics_data <- papers_weighted %>%
+  filter(!is.na(Income.group), !is.na(category)) %>%
+  group_by(Income.group, category) %>%
+  summarise(
+    weighted_n = sum(country_weight, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  rename(Topic = category) %>%
+  mutate(
+    Income.group = factor(
+      Income.group,
+      levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+    ),
+    Topic = factor(
+      Topic,
+      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+    )
+  ) %>%
+  group_by(Income.group) %>%
+  mutate(
+    share = weighted_n / sum(weighted_n),
+    total = sum(weighted_n)        # weighted total "papers" per income group
+  ) %>%
+  ungroup()
+
+## 4. Data for the top "total papers" row -----------------------------------
+
+totals_data <- topics_data %>%
+  group_by(Income.group) %>%
+  summarise(total = round(first(total)), .groups = "drop") %>%
+  mutate(Topic = "Total papers")   # so it can align in the same plot if needed
+
+## 5. Plot (heatmap) --------------------------------------------------------
+
+pdf("plots/Figure2_category_paper_by_income_weight.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = totals_data,
+            aes(x = Income.group, y = "Total papers", fill = total),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = totals_data,
+    aes(x = Income.group, y = "Total papers",
+        label = paste0("n=", total),
+        color = "black"),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Total papers",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(totals_data$total),
+    labels = range(totals_data$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = topics_data,
+            aes(x = Income.group, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = topics_data,
+    aes(x = Income.group, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        # label = paste0(
+        #     scales::percent(share, accuracy = 1), 
+        #     "\n(", n_papers, ")"
+        #   ),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Share of papers\nwithin income group ",
+    low = "#f7f4f9", high = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data$share), max(topics_data$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Income group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y = element_text(size = 22),
+    axis.title = element_text(size = 22),
+    legend.position = "top"
+  )
+dev.off()
+
+```
+
 ## Visualizations by exposure group
@@
 ```
+
+# Visualizations by exposure group, with proportional weighting
+
+```{r visualisation categoryPollution weighting}
+
+#-------------------------------------------------------------------------------
+# papers
+#-------------------------------------------------------------------------------
+
+## 1. Compute country-weights per paper x categoryPollution -----------------------
+
+# For each paper, how many distinct countries are there in total,
+# and how many of those are in each income group?
+paper_country_weights <- papers %>%
+  distinct(ID, country, categoryPollution) %>%  # ignore category here
+  group_by(ID) %>%
+  mutate(n_countries_total = n_distinct(country)) %>%
+  group_by(ID, categoryPollution, n_countries_total) %>%
+  summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
+  mutate(country_weight = n_countries_group / n_countries_total)
+# Example: if a paper has 2 countries (1 low, 1 high) then
+# country_weight = 0.5 for low income and 0.5 for high income.
+
+## 2. Attach weights back to the (paper x categoryPollution x category) rows ------
+
+papers_weighted <- papers %>%
+  left_join(
+    paper_country_weights %>%
+      dplyr::select(ID, categoryPollution, country_weight),
+    by = c("ID", "categoryPollution")
+  )
+
+## 3. Compute weighted category shares within each categoryPollution --------------
+
+topics_data <- papers_weighted %>%
+  filter(!is.na(categoryPollution), !is.na(category)) %>%
+  group_by(categoryPollution, category) %>%
+  summarise(
+    weighted_n = sum(country_weight, na.rm = TRUE),
+    .groups = "drop"
+  ) %>%
+  rename(Topic = category) %>%
+  mutate(
+    Topic = factor(
+      Topic,
+      levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+    )
+  ) %>%
+  group_by(categoryPollution) %>%
+  mutate(
+    share = weighted_n / sum(weighted_n),
+    total = sum(weighted_n)        # weighted total "papers" per categoryPollution
+  ) %>%
+  ungroup()
+
+## 4. Data for the top "total papers" row -----------------------------------
+
+totals_data <- topics_data %>%
+  group_by(categoryPollution) %>%
+  summarise(total = round(first(total)), .groups = "drop") %>%
+  mutate(Topic = "Total papers")   # so it can align in the same plot if needed
+
+## 5. Plot (heatmap) --------------------------------------------------------
+
+pdf("plots/Figure2_category_paper_by_exposure_weight.pdf", width = 12, height = 10)
+ggplot() +
+  # --- top totals row ---
+  geom_tile(data = totals_data,
+            aes(x = categoryPollution, y = "Total papers", fill = total),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = totals_data,
+    aes(x = categoryPollution, y = "Total papers",
+        label = paste0("n=", total),
+        color = "black"),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Total papers",
+    low = "#F6DBA7", high = "#F7B32B",
+    breaks = range(totals_data$total),
+    labels = range(totals_data$total)
+  ) +
+  scale_color_identity() +
+
+  ggnewscale::new_scale_fill() +  # reset fill scale
+
+  # --- main topic rows ---
+  geom_tile(data = topics_data,
+            aes(x = categoryPollution, y = Topic, fill = share),
+            color = "white", linewidth = 0.8) +
+  geom_text(
+    data = topics_data,
+    aes(x = categoryPollution, y = Topic,
+        label = scales::percent(share, accuracy = 1),
+        # label = paste0(
+        #     scales::percent(share, accuracy = 1), 
+        #     "\n(", n_papers, ")"
+        #   ),
+        color = ifelse(share > 0.25, "white", "black")),
+    size = 7
+  ) +
+  scale_fill_gradient(
+    name = "Share of papers\nwithin exposure group ",
+    low = "#f7f4f9", high = "#8B475D",
+    labels = scales::percent,
+    breaks = c(min(topics_data$share), max(topics_data$share))
+  ) +
+  scale_color_identity() +
+  labs(x = "Exposure group", y = NULL) +
+  theme_minimal(base_size = 19) +
+  theme(
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 30, hjust = 1, size = 22),
+    axis.text.y = element_text(size = 22),
+    axis.title = element_text(size = 22),
+    legend.position = "top"
+  )
+dev.off()
+
+```

### 2025-12-09 draftFiguresPaper.Rmd
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP

--- 2025-10-27 draftFiguresPaper.Rmd
@@
        dpi=600, width = 16, height = 8, units='cm')
-
-```
-
-```{r putting data together}
-
-library(tidyverse)
-"%ni%" = Negate("%in%")
-
-datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
-
-datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
-
-datasetCountries <- datasetCountries %>% 
-  select(ID, countryPartialF) %>% 
-  rename(country = countryPartialF)
-
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
-                                        by = "ID") %>% 
-  select(-reader)
-
-datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
-
-table(datasetClassificationCountries$country)
-
-# 1) split into one row per country
-datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ",\\s*") %>%
-  mutate(countries = str_trim(country))
-
-datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
-                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
-                                                               datasetClassificationCountries_long$countries))
-# fix country names 
-library(countrycode)
-
-# 1. Direct mapping
-datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-# 2) get total papers by country
-papers_by_country <- datasetClassificationCountries_long_clean %>%
-  group_by(iso3, .drop = F) %>% 
-  count(iso3, name = "n_papers") %>%
-  arrange(desc(n_papers))
-
-datasetPollution <- read.csv("dataNew/GlobalPM25-V6GL0203-Annual-1998-2023-wThresFrac.csv")
-
-datasetPollution_avg <- datasetPollution %>% 
-  filter(Year %in% 2019) %>%
-  group_by(Region) %>% 
-  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
-
-datasetPollution2 <- read.csv("dataNew/GlobalPM25-V5GL0502-Annual-1998-2023-wThresFrac.csv")
-
-datasetPollution2_avg <- datasetPollution2 %>% 
-  filter(Year %in% 2019) %>%
-  filter(Region %in% c("Hong Kong", "Taiwan")) %>% 
-  group_by(Region) %>% 
-  summarize(pm25Avg = mean(Population.Weighted.PM2.5..ug.m3., na.rm=T))
-
-datasetPollution_avg <- rbind(datasetPollution_avg, datasetPollution2_avg)
-
-datasetPollution_ctr <- datasetPollution_avg %>%
-  mutate(
-    iso3 = countrycode(Region,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(!is.na(iso3)) %>% 
-  rename(countryName = Region)
-
-papers_by_country_pollution <- merge(papers_by_country, datasetPollution_ctr, by = "iso3", all.x = T, all.y = T)
-
-fullCountrylist <- unique(countrycode::codelist$iso3c)
-
-papers_by_country_pollution$n_papers <- ifelse(!is.na(papers_by_country_pollution$n_papers), papers_by_country_pollution$n_papers, 0)
-
-# iso3_un_members <- c(
-#   "AFG","ALB","DZA","AND","AGO","ATG","ARG","ARM","AUS","AUT","AZE",
-#   "BHS","BHR","BGD","BRB","BLR","BEL","BLZ","BEN","BTN","BOL","BIH",
-#   "BWA","BRA","BRN","BGR","BFA","BDI","CPV","KHM","CMR","CAN","CAF",
-#   "TCD","CHL","CHN","COL","COM","COG","COD","CRI","CIV","HRV","CUB",
-#   "CYP","CZE","DNK","DJI","DMA","DOM","ECU","EGY","SLV","GNQ","ERI",
-#   "EST","SWZ","ETH","FJI","FIN","FRA","GAB","GMB","GEO","DEU","GHA",
-#   "GRC","GRD","GTM","GIN","GNB","GUY","HTI","HND","HUN","ISL","IND",
-#   "IDN","IRN","IRQ","IRL","ISR","ITA","JAM","JPN","JOR","KAZ","KEN",
-#   "KIR","PRK","KOR","KWT","KGZ","LAO","LVA","LBN","LSO","LBR","LBY",
-#   "LIE","LTU","LUX","MDG","MWI","MYS","MDV","MLI","MLT","MHL","MRT",
-#   "MUS","MEX","FSM","MDA","MCO","MNG","MNE","MAR","MOZ","MMR","NAM",
-#   "NRU","NPL","NLD","NZL","NIC","NER","NGA","MKD","NOR","OMN","PAK",
-#   "PLW","PAN","PNG","PRY","PER","PHL","POL","PRT","QAT","ROU","RUS",
-#   "RWA","KNA","LCA","VCT","WSM","SMR","STP","SAU","SEN","SRB","SYC",
-#   "SLE","SGP","SVK","SVN","SLB","SOM","ZAF","SSD","ESP","LKA","SDN",
-#   "SUR","SWE","CHE","SYR","TJK","TZA","THA","TLS","TGO","TON","TTO",
-#   "TUN","TUR","TKM","TUV","UGA","UKR","ARE","GBR","USA","URY","UZB",
-#   "VUT","VEN","VNM","YEM","ZMB","ZWE"
-# )
-
-# papers_by_country_pollution_filteredUN <- subset(papers_by_country_pollution, iso3 %in% iso3_un_members) %>% 
-#   filter(!is.na(iso3))
-
-## adding survey nb
-
-surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
-
-surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
-                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
-                                                 surveyClassification$surveyedCountries)
-
-# 1) split into one row per country
-surveyCountries_long <- surveyClassification %>%
-  separate_rows(surveyedCountries, sep = ";\\s*") %>%
-  mutate(countries = str_trim(surveyedCountries))
-
-# fix country names first manual then with iso3
-
-surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
-                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
-                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
-                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
-                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
-                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
-                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
-                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
-
-# 1. Direct mapping
-surveyCountries_long_clean <- surveyCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-# 2) get total surveys by country
-surveys_by_country <- surveyCountries_long_clean %>%
-  filter(Inclusion == 1) %>% 
-  group_by(iso3, .drop = F) %>% 
-  count(iso3, name = "n_surveys") %>%
-  arrange(desc(n_surveys))
-
-papers_by_country_pollution_survey <- merge(papers_by_country_pollution, surveys_by_country, by = "iso3", all.x = T, all.y = T)
-
-papers_by_country_pollution_survey$n_surveys <- ifelse(is.na(papers_by_country_pollution_survey$n_surveys), 0, papers_by_country_pollution_survey$n_surveys)
-
-population <- read.csv("dataNew/GHS_pop.csv", sep = ";")
-
-popCtr <- population %>%
-  group_by(GADM_ISO) %>% 
-  summarize(population2015 = sum(X2015))
-
-summaryDataset <- merge(papers_by_country_pollution_survey, popCtr, by.x = "iso3", by.y = "GADM_ISO", all.x = T, all.y = F)
-
-summaryDataset$categoryPollution <- cut(summaryDataset$pm25Avg, breaks = c(0, 10, 15, 25, 35, 100), 
-                  labels = c("very low", "low", "medium", "high", "very high"), 
-                  right = FALSE)
-
-summaryDataset$isChina <- ifelse(summaryDataset$iso3 == "CHN", 1, 0)
-summaryDataset$isIndia <- ifelse(summaryDataset$iso3 == "IND", 1, 0)
-
-### add income groups to the dataset
-
-incomeGroups <- read.csv("dataNew/worldBankIncomeGroups.csv", sep = ";")
-incomeGroups$Income.group <- ifelse(incomeGroups$Code == "VEN", "Upper middle income", incomeGroups$Income.group)
-
-summaryDataset <- merge(summaryDataset, incomeGroups, by.x = "iso3", by.y = "Code", all.x = T, all.y = F)
-
-plot(summaryDataset %>% 
-       select(Income.group, pm25Avg, n_papers, n_surveys, population2015, categoryPollution))
-
-## add region from natural earth
-
-library(rnaturalearth)
-library(terra)
-
-world_countries <- vect(ne_countries(scale = "large", returnclass = "sf"))
-
-world_countries_df <- as.data.frame(world_countries)
-
-world_countries_df$iso_a3 <- ifelse(world_countries_df$iso_a3_eh == "FRA", "FRA",
-                                    ifelse(world_countries_df$iso_a3_eh == "NOR", "NOR", world_countries_df$iso_a3))
-
-summaryDataset2 <- merge(summaryDataset, world_countries_df %>% 
-                          select(continent, region_un, iso_a3), by.x = "iso3", by.y = "iso_a3", all.x = T, all.y = F)
-
-write.csv(summaryDataset2, "output/datasetPlots.csv", row.names = F)
-
-```
-
-```{r long classification dataset - paper}
-
-library(tidyverse)
-library(countrycode)
-
-"%ni%" = Negate("%in%")
-
-datasetClassification <- read.csv("output/dataClassifToPloto4mini.csv")
-
-datasetCountries <- read.csv("output/fullClassifCountryo4mini.csv")
-
-datasetCountries <- datasetCountries %>% 
-  dplyr::select(ID, countryPartialF) %>% 
-  rename(country = countryPartialF)
-
-datasetClassificationCountries <- merge(datasetClassification, datasetCountries,
-                                        by = "ID") %>% 
-  dplyr::select(-reader)
-
-datasetClassificationCountries$country <- ifelse(datasetClassificationCountries$country == "None", "Unknown", datasetClassificationCountries$country)
-
-table(datasetClassificationCountries$country)
-
-# 1) split into one row per country
-datasetClassificationCountries_long <- datasetClassificationCountries %>%
-  separate_rows(country, sep = ",\\s*") %>%
-  mutate(countries = str_trim(country))
-
-datasetClassificationCountries_long$countries <- ifelse(datasetClassificationCountries_long$countries == "Turkyie", "Turkey",
-                                                        ifelse(datasetClassificationCountries_long$countries %in% c("Wales", "England"), "United Kingdom",
-                                                               datasetClassificationCountries_long$countries))
-# fix country names 
-# 1. Direct mapping
-datasetClassificationCountries_long_clean <- datasetClassificationCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-# add the year
-datasetYear <- read.csv("dataNew/datasetToClassifyGPT.csv") %>% 
-  dplyr::select(ID, Year)
-
-datasetClassificationCountries_long_clean <- merge(datasetClassificationCountries_long_clean,
-                                                   datasetYear,
-                                                   by = "ID", all.x = T, all.y = F)
-
-df_cat_long <- datasetClassificationCountries_long_clean %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  dplyr::select(-flag, -None)
-
-write.csv(df_cat_long, "output/paperClassifYear.csv", row.names = F)
-
-```
-
-```{r long classification dataset - survey}
-
-library(tidyverse)
-library(countrycode)
-
-"%ni%" = Negate("%in%")
-
-surveyClassification <- read.csv("dataNew/surveyClassif.csv", sep = ";")
-
-surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Kazakhstan; Kyrgyzstan, Tajikistan, Uzbekistan, Turkmenistan",
-                                                 "Kazakhstan; Kyrgyzstan; Tajikistan; Uzbekistan; Turkmenistan",
-                                                 surveyClassification$surveyedCountries)
-
-# 1) split into one row per country
-surveyCountries_long <- surveyClassification %>%
-  separate_rows(surveyedCountries, sep = ";\\s*") %>%
-  mutate(countries = str_trim(surveyedCountries))
-
-# fix country names first manual then with iso3
-
-surveyCountries_long$countries <- ifelse(surveyCountries_long$countries %in% c("Phillipines", "Philliphines", "Phillippines"), "Philippines",
-                                         ifelse(surveyCountries_long$countries %in% c("Northern Ireland"), "Ireland",
-                                                ifelse(surveyCountries_long$countries %in% c("Czech"), "Czechia",
-                                                       ifelse(surveyCountries_long$countries %in% c("Uruguah"), "Uruguay",
-                                                              ifelse(surveyCountries_long$countries %in% c("Natherlands"), "Netherlands",
-                                                                     ifelse(surveyCountries_long$countries %in% c("Croaatia"), "Croatia",
-                                                                            ifelse(surveyCountries_long$countries %in% c("Kyrgysztan"), "Kyrgyzstan",
-                                                                                   ifelse(surveyCountries_long$countries %in% c("El Savador"), "El Salvador", surveyCountries_long$countries))))))))
-
-# 1. Direct mapping
-surveyCountries_long_clean <- surveyCountries_long %>%
-  mutate(
-    iso3 = countrycode(countries,
-                       origin      = "country.name",
-                       destination = "iso3c")
-  ) %>% 
-  filter(countries %ni% c("", " ", "Kosovo", "Southeast Asia"))
-
-df_cat_long <- surveyCountries_long_clean %>%
-  pivot_longer(
-    cols = c(Perceptions, Priority, Policy, Health, Behavior),
-    names_to  = "category",
-    values_to = "flag"
-  ) %>%
-  # keep only “paper belongs to this category”
-  filter(flag == 1) %>%
-  dplyr::select(-flag)
-
-write.csv(df_cat_long, "output/surveyClassifYear.csv", row.names = F)
 

### 2025-12-17 dataPrep.Rmd
more surveys need to add whether barometer

@@
 
+surveyClassification$surveyedCountries <- ifelse(surveyClassification$surveyedCountries == "Belgium; Denmark; Germany; Spain, Greece; France; Ireland; Italy; Luxembourg; Netherlands; Portugal; United Kingdom",
+                                                 "Belgium; Denmark; Germany; Spain; Greece; France; Ireland; Italy; Luxembourg; Netherlands; Portugal; United Kingdom",
+                                                 surveyClassification$surveyedCountries)
+
 # 1) split into one row per country
@@
 surveyCountries_long_clean_yearly <- surveyCountries_long_clean
+
+table(surveyCountries_long_clean_yearly$year)
+
 surveyCountries_long_clean_yearly$year <- ifelse(surveyCountries_long_clean_yearly$year == "2017 - 2021", 2019, surveyCountries_long_clean_yearly$year)
@@
 ```
+
+```{r count eurobarometer}
+
+library(tidyverse)
+surveys <- read.csv("output/surveyClassifYear.csv")
+
+surveys <- surveys %>% 
+  filter(Inclusion == 1)
+
+surveys <- surveys %>% 
+  filter(!is.na(iso3))
+
+surveys <- surveys %>% 
+  filter(!(iso3 == "FRA" & continent == "Seven seas (open ocean)")) %>% 
+  mutate(item_id = paste0(iso3, id, year)) %>%                      # unique survey item ID
+  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
+  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
+
+eurobarometer_surveys <- surveyClassification %>%
+  filter(grepl("Eurobarometer", SurveyName, ignore.case = TRUE))
+
+nrow(eurobarometer_surveys)
+
+```

### 2025-12-17 draftFigure1.Rmd
fix bug year survey item
adding weighted category paper count also remove duplicate rows in dataset for FRA and CYP
more surveys need to add whether barometer

@@
 
+```{r panels with eurobarometer highlight}
+library(tidyverse)
+dataset <- read.csv("output/datasetPlots.csv")
+
+```
+
 ```{r maps}
@@
 
-my_breaks = c(1, 5, 10, 20)
+my_breaks = c(1, 5, 10, 30, 110)
 
@@
     colours = c("#F3EBF8", "#D4ABEF", "#541F78"),
-    name    = "Papers",
+    name    = "Survey items",
     trans   = "log10",

### 2025-12-17 dataPrep.Rmd
updated code and figure but need to review fig2 and update exposure fig2 heatmap

@@
 
+# count separately for eurobarometer and non-eurobarometer
+surveyCountries_long_clean_t_euro <- surveyCountries_long_clean_t %>% 
+  mutate(eurobarometerSource = ifelse(grepl("Eurobarometer", source, ignore.case = TRUE), 1, 0))
+
 # 2) get total surveys by country
@@
 
-```
+# create a barometer-split dataset
 
-```{r count eurobarometer}
+surveyCountries_long_clean_t_euro <- surveyCountries_long_clean_t_euro
 
-library(tidyverse)
-surveys <- read.csv("output/surveyClassifYear.csv")
+surveys_by_country_euro <- surveyCountries_long_clean_t_euro %>%
+  filter(Inclusion == 1) %>% 
+  group_by(iso3, eurobarometerSource, .drop = F) %>% 
+  count(iso3, name = "n_surveys") %>%
+  arrange(desc(n_surveys))
 
-surveys <- surveys %>% 
-  filter(Inclusion == 1)
+summaryDataset_euro <- merge(summaryDataset2 %>% 
+                               select(-n_surveys) %>% 
+                               crossing(eurobarometerSource = c(0, 1)),
+                             surveys_by_country_euro,
+                             by = c("iso3", "eurobarometerSource"), all.x = T, all.y = T)
 
-surveys <- surveys %>% 
-  filter(!is.na(iso3))
+summaryDataset_euro$n_surveys <- ifelse(is.na(summaryDataset_euro$n_surveys), 0, summaryDataset_euro$n_surveys)
 
-surveys <- surveys %>% 
-  filter(!(iso3 == "FRA" & continent == "Seven seas (open ocean)")) %>% 
-  mutate(item_id = paste0(iso3, id, year)) %>%                      # unique survey item ID
-  distinct(item_id, categoryPollution, category) %>%          # each item counts once per (exposure, topic)
-  count(categoryPollution, category, name = "n_items") %>%    # number of items mentioning category
-
-eurobarometer_surveys <- surveyClassification %>%
-  filter(grepl("Eurobarometer", SurveyName, ignore.case = TRUE))
-
-nrow(eurobarometer_surveys)
+write.csv(summaryDataset_euro, "output/datasetPlots_eurobarometer.csv", row.names = F)
 

### 2025-12-17 draftFigure1.Rmd
updated code and figure but need to review fig2 and update exposure fig2 heatmap

@@
   theme_minimal() +
-  labs(y = "Total number of survey items", x = "income Group") +
+  labs(y = "Total number of survey items", x = "Income group") +
   theme(legend.position = "none") +
@@
 ```{r panels with eurobarometer highlight}
+# globally there are 2390 survey items from eurobarometer out of 3140 total
+
 library(tidyverse)
-dataset <- read.csv("output/datasetPlots.csv")
+dataset <- read.csv("output/datasetPlots_eurobarometer.csv")
+
+dataset <- dataset %>% 
+  mutate(isDiff = ifelse(eurobarometerSource == 1, "Eurobarometer", "Rest"))
+
+summarySurveysInc <- dataset %>% 
+  group_by(Income.group, isDiff) %>% 
+  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
+  filter(!is.na(Income.group))
+
+summarySurveysInc$Income.group <- factor(
+  summarySurveysInc$Income.group,
+  levels = rev(c(
+    "High income",
+    "Upper middle income",
+    "Lower middle income",
+    "Low income"
+  )),
+  ordered = TRUE
+)
+
+totals <- summarySurveysInc %>%
+  group_by(Income.group) %>%
+  summarise(total = sum(nTotal), .groups = "drop")
+
+p2 <- summarySurveysInc %>%
+  ggplot(aes(x = Income.group, y = nTotal)) +
+  geom_bar(stat = "identity", aes(fill = isDiff)) +
+  geom_text(
+    data = totals,
+    aes(x = Income.group, y = total, label = total),
+    vjust = -0.3, fontface = "bold"
+  ) +
+  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
+  theme_minimal() +
+  labs(y = "Total number of survey items", x = "Income group") +
+  theme(legend.position = "none") +
+  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping
+
+p2
+
+pdf("plots/Figure1_surveys_by_income_eurob.pdf", width = 10)
+p2
+dev.off()
+
+## then surveys per exposure
+
+summarySurveysExp <- dataset %>% 
+  group_by(categoryPollution, isDiff) %>% 
+  summarize(nTotal = sum(n_surveys, na.rm = T)) %>% 
+  filter(!is.na(categoryPollution))
+
+summarySurveysExp$categoryPollution <- ifelse(
+  !is.na(summarySurveysExp$categoryPollution),
+  paste0(toupper(substr(summarySurveysExp$categoryPollution, 1, 1)),
+         substr(summarySurveysExp$categoryPollution, 2, nchar(summarySurveysExp$categoryPollution))),
+  NA
+)
+
+summarySurveysExp$categoryPollution <- factor(
+  summarySurveysExp$categoryPollution,
+  levels = rev(c(
+    "Very high",
+    "High",
+    "Medium",
+    "Low",
+    "Very low"
+  )),
+  ordered = TRUE
+)
+
+totals <- summarySurveysExp %>%
+  group_by(categoryPollution) %>%
+  summarise(total = sum(nTotal), .groups = "drop")
+
+p4 <- summarySurveysExp %>%
+  ggplot(aes(x = categoryPollution, y = nTotal)) +
+  geom_bar(stat = "identity", aes(fill = isDiff)) +
+  geom_text(
+    data = totals,
+    aes(x = categoryPollution, y = total, label = total),
+    vjust = -0.3, fontface = "bold"
+  ) +
+  scale_fill_manual(values = c("#E6D0F4", "#B873E7")) +
+  theme_minimal() +
+  labs(y = "Total number of survey items", x = "Pollution exposure") +
+  theme(legend.position = "none") +
+  expand_limits(y = max(totals$total) * 1.05)  # avoid clipping
+
+p4
+
+pdf("plots/Figure1_surveys_by_exposure_eurob.pdf", width = 10)
+p4
+dev.off()
 

### 2025-12-17 draftFigure2.Rmd
edit survey classif with more items
more surveys need to add whether barometer
updated code and figure but need to review fig2 and update exposure fig2 heatmap

@@
 ## subset survey without Flashbarometer
-surveys_sub <- surveys[which(surveys$source != "Flash Eurobarometer 360: Attitudes of Europeans towards air quality"),]
+# surveys_sub <- surveys[which(surveys$source != "Flash Eurobarometer 360: Attitudes of Europeans towards air quality"),]
 
@@
     "None"        = "#F4A460")) +
-  xlim(0, 1850) +
+  xlim(0, 2100) +
   theme_minimal() +
@@
 
-# For each paper, how many distinct countries are there in total,
-# and how many of those are in each income group?
+## Alternative approach: also account for multiple categories per paper -----
 paper_country_weights <- papers %>%
-  distinct(ID, country, Income.group) %>%  # ignore category here
+  distinct(ID, country, Income.group) %>%
   group_by(ID) %>%
@@
   summarise(n_countries_group = n_distinct(country), .groups = "drop") %>%
-  mutate(country_weight = n_countries_group / n_countries_total)
-# Example: if a paper has 2 countries (1 low, 1 high) then
-# country_weight = 0.5 for low income and 0.5 for high income.
-
-## 2. Attach weights back to the (paper x income group x category) rows ------
-
-papers_weighted <- papers %>%
-  left_join(
-    paper_country_weights %>%
-      dplyr::select(ID, Income.group, country_weight),
-    by = c("ID", "Income.group")
-  )
-
-## 3. Compute weighted category shares within each income group --------------
-
-topics_data <- papers_weighted %>%
+  mutate(country_weight = n_countries_group / n_countries_total) %>%
+  dplyr::select(ID, Income.group, country_weight)
+
+papers_pic <- papers %>%
   filter(!is.na(Income.group), !is.na(category)) %>%
+  distinct(ID, Income.group, category) %>%
+  group_by(ID) %>%
+  mutate(n_cat = n_distinct(category)) %>%
+  ungroup() %>%
+  left_join(paper_country_weights, by = c("ID", "Income.group")) %>%
+  mutate(weight = country_weight / n_cat)
+
+topics_data <- papers_pic %>%
   group_by(Income.group, category) %>%
-  summarise(
-    weighted_n = sum(country_weight, na.rm = TRUE),
-    .groups = "drop"
+  summarise(weighted_n = sum(weight, na.rm = TRUE), .groups = "drop") %>%
+  rename(Topic = category) %>%
+  group_by(Income.group) %>%
+  mutate(
+    share = weighted_n / sum(weighted_n),
+    total = sum(weighted_n)
   ) %>%
-  rename(Topic = category) %>%
+  ungroup() %>%
   mutate(
@@
     )
-  ) %>%
-  group_by(Income.group) %>%
-  mutate(
-    share = weighted_n / sum(weighted_n),
-    total = sum(weighted_n)        # weighted total "papers" per income group
-  ) %>%
-  ungroup()
+  )
+
+## 3. Compute weighted category shares within each income group --------------
+
+# topics_data <- papers_weighted %>%
+#   filter(!is.na(Income.group), !is.na(category)) %>%
+#   group_by(Income.group, category) %>%
+#   summarise(
+#     weighted_n = sum(country_weight, na.rm = TRUE),
+#     .groups = "drop"
+#   ) %>%
+#   rename(Topic = category) %>%
+#   mutate(
+#     Income.group = factor(
+#       Income.group,
+#       levels = c("Low income", "Lower middle income", "Upper middle income", "High income")
+#     ),
+#     Topic = factor(
+#       Topic,
+#       levels = rev(c("Perceptions", "Behavior", "Priority", "Policy", "Health"))
+#     )
+#   ) %>%
+#   group_by(Income.group) %>%
+#   mutate(
+#     share = weighted_n / sum(weighted_n),
+#     total = sum(weighted_n)        # weighted total "papers" per income group
+#   ) %>%
+#   ungroup()
 
